---
title: "Myositis"
format:
  html: default
  pdf: default
  docx: default
editor: visual
execute:
  echo: false
editor_options: 
  chunk_output_type: inline
always_allow_html: yes
---

## Reading the data

```{r}
#| message: false
#| label: loading-libraries-and-data

rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")

options(scipen = 999)

# Loading required libraries
library(tidyverse); library(lubridate);library(here)
library(furrr); library(jsonlite);library(arrow)
library(feather)
library(udpipe)
library(tidytext)
library(stringi)
library(arrow)

source(here("config.R"))

# Loading Reddit parquet data

df_start <- arrow::read_feather(file.path(DATA_PATH, "data", "myositis_submissions_incl_subreddits_2005_july2025.feather")) 

df <- df_start 

```

Before cleaning, the data contained `r scales::comma(nrow(df_start))` Reddit posts, posted by `r scales::comma(length(unique(df_start$author)))` authors, posted in `r scales::comma(length(unique(df_start$subreddit)))` subreddits. `r scales::comma( sum(df_start$category == "myositis"))` posts were related to Myositis, `r scales::comma( sum(df_start$category == "muscular dystrophy"))` were related to Muscular Dystrophy, and `r scales::comma(sum(df_start$category == "mixed"))` posts referred to both.

Several steps were taken to clean the data. The following posts were removed:

-   posts selected based on myopathy/ies as a keyword but not referring to inflammatory of necrotizing myopathy/ies
-   duplicate posts (identical text fields (title and selftext) and posted within 14 days of each other by the same author)
-   deleted selftext
-   posts written by by accounts that are clearly bots
-   non English posts
-   posts containing 30 or more asterisks or 10 or more urls's
-   posts with fewer than 10 words
-   posts containing the word "product name"
-   posts from subreddits that were about animals
-   posts from subreddits or authors that were not about personal experiences or questions
-   personal subreddits (u/)

```{r}
#| message: false
#| label: select correct combinations of myopathy + myositis related words

#library(stringr)
#library(dplyr)
#library(tidyr) # For unnest()

# Word stems to allow variations
myo_keywords <- c("necrotizing myopathy", "necrotizing myopathies", "necrotising myopathy", "necrotising myopathies", 
  "inflammatory myopathy","inflammatory myopathies",
  "myositis", "polymyositis", "dermatomyositis", "synthetase syndrome", "synthetasesyndrome", "anti-jo-1", "antijo1", "anti jo 1"
)

# Escape special characters and add \\w* to allow plural/suffix variations
myo_keywords <- str_replace_all(myo_keywords, "[-()]", "\\\\$0")
myo_keywords <- paste0(myo_keywords, "\\w*")

# regex pattern construction
keyword_pattern <- paste(myo_keywords, collapse = "|")

# Extract keyword occurrences from the 'text' column
words_overview_text <- df |>
  mutate(matches = str_extract_all(text, regex(keyword_pattern, ignore_case = TRUE))) |>
  unnest(matches) |>
  mutate(matches = tolower(matches)) |>
  count(matches, sort = TRUE)

# Extract keyword occurrences from the 'subreddit' column
words_overview_subreddit <- df |>
  mutate(matches = str_extract_all(subreddit, regex(keyword_pattern, ignore_case = TRUE))) |>
  unnest(matches) |>
  mutate(matches = tolower(matches)) |>
  count(matches, sort = TRUE)

df_excluded <- df |>
  filter(
    # The '!' negates the entire condition inside the parentheses
    ! (
      str_detect(text, regex(keyword_pattern, ignore_case = TRUE)) |
      str_detect(subreddit, regex(keyword_pattern, ignore_case = TRUE))
    )
  )

myopath_excluded <- nrow(df_excluded)

# Filter original df to include only rows with a match in either 'text' or 'subreddit'
df <- df |> 
  filter(str_detect(text, regex(keyword_pattern, ignore_case = TRUE)) | 
         str_detect(subreddit, regex(keyword_pattern, ignore_case = TRUE)))

#remove tension myositis and other non inflanmtory types of myositis
n_before <- nrow(df)

df <- df |>
  filter(
    !str_detect(
      paste(subreddit, text, sep = " "),
      regex("tension\\s*myositis(?:syndrome)?", ignore_case = TRUE)
    ) &
    !str_detect(text, regex("pyomyositis", ignore_case = TRUE)) &
    !str_detect(text, regex("myositis ossificans", ignore_case = TRUE)) &
    !str_detect(text, regex("orbital myositis", ignore_case = TRUE))
  )

n_after <- nrow(df)
n_removed <- n_before - n_after

cat("Rows before:", scales::comma(n_before),
    "after:", scales::comma(n_after),
    "removed:", scales::comma(n_before - n_after), "\n")

```

Since myopathy is a broad term that also refers to non-inflammatory muscle conditions, in a first cleaning step, we excluded posts that did not explicitly refer to necrotizing or inflammatory myopathy/ies (n = `r scales::comma(myopath_excluded)`), while keeping all other posts. Also, posts referring to tension myositis, orbital myositis and myositis ossificans all of which are outside the scope of this study, were excluded (n = `r scales::comma(n_removed)`), leaving `r scales::comma(nrow(df))` posts.

```{r}
#| message: false
#| label: (un)select subreddits

exclude_values_general <- c("Step1","Step2","scholar","paidstudies","PaidStudyHub","paidstudy","autotldr","noveltranslations","TheDahmerCase","Dahmer","conspiracy","conspiracyNOPOL","conspiracyundone","sarasota","googletranslate","medicalschool","takecareofmayaFree","skulduggerypleasant","remotedaily","motivation","NBME","zmarter","ChinaInnovation","ketoscience","COVIDZero","Vitiligo","FuckTravisScott","PSSD","buildapc","DebateVaccine","VaxTalk","Autoimmunefaq","SupplementsGuide","weedstocks","medicalschoolanki","anki","dirtypenpals","lakers","OnePiece","tabled","MongoloidCJ","PackItUpinfo","remotedaily","Watches","thesopranos","funny","Cardiomyopathy", "manywords",
                            "CABAinvest", "wallstreetbets")

exclude_values_animals <- c("AskVet","dogs","puppy101","vet","veterinary","Greyhounds","Petloss","CaneCorso","VetTech","veterinarian","Veterinary","GreatDane","Pets","goldenretrievers","u_bobbiesocksthepug","roughcollies","BelgianMalinois","aww","germanshepherds","pitbulls","dogs","cats","Rabbits","CatAdvice","Boxer","DobermanPinscher","AdviceAnimals","Equestrian","greatdanes","CATHELP","blackcats","bengalcats","catfood","SiberianCats","standardissuecat","catcare","RenalCats","Siamesecats","TuxedoCats","IllegallySmolCats","Catahoula","Feral_Cats","catpics","SupermodelCats","petyoloblog","BeardedDragons","BeardedDragonHelp","BeardedDragon","bengalcatfacts","bengals","bernesemountaindog","bernesemountaindogs","BorderCollie","BorderTerrier","BostonTerrier","BoxerDogs","Boxers","britishshorthair","BritishShorthaircats","Horses","BirdHealth","Chihuahua","PetAdvice","akita","corgi","Pitbull", "National_Pet_Adoption", "greatpyrenees", "ridgebacks", "AfricanGrey")

unwanted_authors <- c("myatomsareyouratoms","Asher_potter","D_Upadhaya","Mentioned_Videos","hibernating_brain","Mk_Makanaki","subredditreports","-NervousPudding-","0100001001010011","Firm-Twist-4497","MuscularDystrophy-ModTeam","Rabbits-ModTeam","SlightlyMadAngus","Lupinepublishers-ACR", "mushroomlover345", "Gir0", "FrozenPixel", "namenerd400")

excluded_subreddits <- tolower(c(exclude_values_general, exclude_values_animals))

excluded_subreddits <- tolower(c(exclude_values_general, exclude_values_animals))

n_before <- nrow(df)

df <- df |>
  mutate(
    subreddit_lc = tolower(subreddit),
    title0 = coalesce(title, ""),
    selftext0 = coalesce(selftext, "")
  ) |>
  filter(
    !(
      subreddit_lc %in% excluded_subreddits |
      str_detect(subreddit, regex("penny", ignore_case = TRUE)) |
      str_detect(subreddit, regex("dog", ignore_case = TRUE)) |
      str_detect(subreddit, regex("^u_", ignore_case = TRUE)) |
      author %in% unwanted_authors |
      str_detect(title0, regex("masticatory", ignore_case = TRUE)) |
      str_detect(selftext0, regex("masticatory", ignore_case = TRUE))
    )
  ) |>
  select(-subreddit_lc, -title0, -selftext0)

#remove bots
df <- df |>
  filter(!str_detect(author, regex("bot$", ignore_case = TRUE))) |>
  filter(!str_detect(author, regex("^auto", ignore_case = TRUE))) |>
  filter(!str_detect(author, regex("admin$", ignore_case = TRUE))) 

n_after <- nrow(df)
n_removed <- n_before - n_after 

df <- df |>
  filter(
    !str_detect(text, regex("dahmer", ignore_case = TRUE)) 
  )

cat("Rows before:", scales::comma(n_before),
    "after:", scales::comma(n_after),
    "removed:", scales::comma(n_before - n_after), "\n")
```

The next cleaning step consisted of (a.o.) removing posts from subreddits which were about animals (which can also develop myositis), subreddits clearly not directly related to personal experiences with myositis (e.g., subreddits about serial killer Jeffrey Dahmer, a polymyositis patient), and personal subreddits (u/), thus excluding `r n_removed` posts, leaving `r scales::comma(nrow(df))` messages.

```{r}
#| message: false
#| label: (un)select deleted posts and duplicates

n_before <- nrow(df)

# Remove [removed]/[deleted] or empty title/selftext 

df <- df |>
  mutate(
    across(c(title, selftext), ~ str_trim(.x)),
    across(c(title, selftext), ~ na_if(.x, ""))   # convert "" -> NA
  ) |>
  filter(
    !if_any(
      c(title, selftext),
      ~ is.na(.x) |
        str_detect(.x, regex("^\\s*\\[(removed|deleted)\\]\\s*$", ignore_case = TRUE))
    )
  )


# Flag and drop duplicates within 14 days, per author ----

df <- df |>
  mutate(
    created_utc = as.POSIXct(created_utc, origin = "1970-01-01"),
    rowid = row_number()
  )

dup_flag <- function(dat, col) {
  col_sym <- sym(col)
  
  dat |>
    filter(!is.na(!!col_sym)) |>
    group_by(author, !!col_sym) |>
    arrange(created_utc, .by_group = TRUE) |>
    mutate(
      prev = lag(created_utc),
      .dup = !is.na(prev) & as.numeric(difftime(created_utc, prev, units = "days")) <= 14
    ) |>
    ungroup() |>
    transmute(rowid, !!paste0(col, "_dup") := .dup)
}

cols <- intersect(c("title", "selftext"), names(df))

flags <- purrr::map(cols, ~ dup_flag(df, .x)) |>
  reduce(full_join, by = "rowid")

df <- df |>
  left_join(flags, by = "rowid") |>
  mutate(
    across(ends_with("_dup"), ~ coalesce(.x, FALSE)),
    to_remove = if_any(ends_with("_dup"), identity)
  ) |>
  filter(!to_remove) |>
  select(-rowid, -ends_with("_dup"))

n_after <- nrow(df)
n_removed <- n_before - n_after

cat("Rows before:", scales::comma(n_before),
    "after:", scales::comma(n_after),
    "removed:", scales::comma(n_removed), "\n")


```

Next, posts where removed the title or selftext was empty or consisted only of the markers \[removed\] or \[deleted\], and we excluded potential duplicate posts by the same author. Posts were considered duplicates if they had identical title or selftext content and were submitted within 14 days of each other. Only the earliest post in each duplicate set was retained; later duplicates were excluded. Through these steps, `r scales::comma(n_removed)` posts were excluded, leaving `r scales::comma(nrow(df))` posts.

```{r}
#| message: false
#| label: remove posts with many urls's, asterisks, line breaks

n_before <- nrow(df)

df <- df |>
  mutate(
    # Count URLs
    n_urls_title = str_count(title, pattern = "https?://\\S+|www\\.\\S+"),
    n_urls_selftext = str_count(selftext, pattern = "https?://\\S+|www\\.\\S+"),

    # Count asterisks
    n_asterisks_title = str_count(title, fixed("*")),
    n_asterisks_selftext = str_count(selftext, fixed("*"))
  ) |>
  # Replace NA counts with 0
  replace_na(list(
    n_urls_title = 0, n_urls_selftext = 0, 
    n_asterisks_title = 0, n_asterisks_selftext = 0
  )) |>
  # Add totals per row
  mutate(
    n_urls_total = n_urls_title + n_urls_selftext,
    n_asterisks_total = n_asterisks_title + n_asterisks_selftext
  )

df <- df |>
  filter(
   n_asterisks_total < 11, 
   n_urls_total < 5,
  is.na(selftext) | !str_detect(selftext, regex("product name", ignore_case = TRUE))
)

library(purrr)

df <- df |>
  mutate(
    # count line breaks
    n_lines = str_count(selftext, "\n"),

    # split into lines
    lines = str_split(selftext, "\n"),

    # count words per line
    line_lengths = map(lines, ~ str_count(.x, "\\S+")),

    # proportion of lines with < 5 words
    prop_short = map_dbl(line_lengths, ~ mean(.x < 5, na.rm = TRUE)),

    # define: list-like if many breaks AND mostly short lines
    is_list_like = n_lines >= 20 & prop_short >= 0.7
  )

df <- df |> filter(!is_list_like)

df |> summarise(
  over_10  = sum(n_lines > 10),
  over_20  = sum(n_lines > 20),
  over_50  = sum(n_lines > 50),
  over_100 = sum(n_lines > 100),
  over_200 = sum(n_lines > 200)
)

# filter out 100% only short sentences & >= 50 sentences
df <- df |>
  filter(prop_short != 1 | n_lines >= 50) 
n_after <- nrow(df)
n_removed <- n_before - n_after 

cat("Rows before:", scales::comma(n_before),
    "after:", scales::comma(n_after),
    "removed:", scales::comma(n_before - n_after), "\n")
```

In a next step, we aimed to remove posts that look like spam, commercial content, or non-discursive lists. Posts containing excessive URLs, formatting markers, or explicit promotional cues were removed, as were submissions dominated by long sequences of short, list-like lines (e.g., symptom or supplement lists), which do not provide interpretable patient perspectives. Through quantifying line breaks and the proportion of very short sentences, we identified and excluded content that lacked coherent narrative structure. During this step, `r scales::comma(n_removed)` were removed, leaving `r scales::comma(nrow(df))` posts

```{r}
#| message: false
#| label: removing non English selftext and text < 10 words

n_before <- nrow(df)

df <- df |>
  filter(
    n_words > 9, # only posts with 10 words or more
    cld2::detect_language(text) == "en",                              # keep English
    !is.na(subreddit), subreddit != "omoshiroi_eigo_bunpou",    # drop non-English sub
      )

n_after <- nrow(df)
n_removed <- n_before - n_after 

cat("Rows before:", scales::comma(n_before),
    "after:", scales::comma(n_after),
    "removed:", scales::comma(n_before - n_after), "\n")

subr <- df |> 
  count(subreddit, name = "freq") |>          # frequency
  arrange(desc(freq)) |>                      # sort
  mutate(
    prop = freq / sum(freq),                  # proportion
    cum_prop = cumsum(prop)                   # cumulative proportion
  )
length(unique(df$author))
```

In the next step, all non-English posts and posts with fewer than 10 words were excluded, after which `r scales::comma(nrow(df))` posts remained, posted by `r scales::comma(length(unique(df$author)))` unique authors. The figure below shows the number of myositis related posts per month.

```{r}
#| message: false
#| label: create lemmatized and cleaned df$text_clean column with udpipe

download_info <- udpipe_download_model(language = "english-ewt")

# The 'download_info' object (a data frame) contains the path to the downloaded model.
# We'll extract the actual filename/path from it.
model_path_downloaded <- download_info$file_model

# Check if the model file exists (it should, after the download attempt)
if (!file.exists(model_path_downloaded)) {
  stop("Failed to download the English udpipe model.")
}

# Load the English udpipe model into your R session
ud_model <- udpipe_load_model(file = model_path_downloaded)

# 1) Light cleaning + Unicode normalization (keep punctuation for UDPipe)
df$text_clean <- df$text |>
  str_replace_all("https?://\\S+|www\\.\\S+", " ") |>
  stri_trans_general("NFKC") |>                 # normalize accents/compatibility
  str_replace_all("\\p{Cf}", " ") |>            # remove zero-width/format chars
  str_squish()

# 2) Annotate

# make sure df has a doc_id column
if (!"doc_id" %in% names(df)) {
  df$doc_id <- as.character(seq_len(nrow(df)))
}

x <- udpipe_annotate(ud_model, x = df$text_clean, doc_id = df$doc_id)
x_df <- as.data.frame(x) |>
  mutate(lemma = ifelse(tolower(token) == "feedback", "feedback", lemma))

# 3) Post-process: keep content words; fall back to token if tag looks wrong
content_upos <- c("NOUN","PROPN","ADJ","VERB")

domain_map <- c( # protect key RA/biomed terms
  "joint"="joint", "joints"="joint",
  "flare"="flare", "flares"="flare",
  "ra"="ra", "ccp"="ccp", "rf"="rf",
  "myositis"="myositis", "neuropathy"="neuropathy"
)

lemmas_filtered <- x_df |>
  transmute(
    doc_id, token = tolower(token), lemma = tolower(lemma), upos
  ) |>
  # If UDPipe calls a long alphabetic token a PRON/DET/ADP, treat lemma as the token
  mutate(bad_tag = (!upos %in% content_upos) & grepl("^[a-z][a-z'-]{3,}$", token),
         lemma = if_else(bad_tag, token, lemma)) |>
  # Sanity: if lemma has weird chars, fall back to token
  mutate(lemma = if_else(!grepl("^[a-z][a-z'-]*$", lemma), token, lemma)) |>
  # Domain dictionary overrides
  mutate(lemma = if_else(token %in% names(domain_map), domain_map[token], lemma)) |>
  # Finally filter to content words and non-stopwords (but keep noun "back")
  filter(upos %in% content_upos) |>
  filter((lemma == "back" & upos == "NOUN") | !(lemma %in% stop_words$word))

lemmas_filtered <- lemmas_filtered |>
  filter(!(lemma %in% c("ive", "im", "dont", "cant", "wont"))) |>
  mutate(lemma = if_else(lemma == "jous", "joint", lemma)) #correcting a mistake in the udpipe process which replaces joint by jous

# 4) Rebuild lemmatized text
lemmatized_text_df <- lemmas_filtered |>
  group_by(doc_id) |>
  summarise(lemmatized_text = paste(lemma, collapse = " "), .groups="drop")

df <- df |> left_join(lemmatized_text_df, by = "doc_id")
# You can inspect the new column
head(df |> select(doc_id, text_clean, lemmatized_text))

#write_feather(df, file.path(DATA_PATH, "data", "myositis_submissions_2005_july2025_clean.feather"))
#write_feather(x_df, file.path(DATA_PATH, "data", "myositis_lemmatized_udpipe.feather"))
```

```{r}
df_summary <- df |>
  mutate(year_month = as.Date(floor_date(created_utc, "month"))) |>
  group_by(year_month) |>
  summarize(count = n())

p <- ggplot(df_summary, aes(x = year_month, y = count)) +
  geom_point() +
  geom_jitter() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(title = "Number of Mentions per Month",
       x = "Year",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))

p
```

The number of myositis related posts has increased to close to 100 per month in 2024, showing a steep growth after 2020 (as did Reddit as a whole). Over half of the submissions were posted in the years 2024 or 2025, as the table below shows.

```{r}
#| message: false
#| echo: false

library(knitr)
library(kableExtra)
subr <- df |>
  count(year) |>
  rename(freq = n) |>
  mutate(percentage = (freq / sum(freq)) * 100)|>
  arrange(desc(year)) |>
  mutate(cum_percentage = cumsum(percentage)) |>
  mutate(percentage = round(percentage, 1),
         cum_percentage = round(cum_percentage,1))

# Create table
subr |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

Mentions of myositis are spread across a wide range of subreddits, as the tables below show. There are subreddits specifically for myositis (e.g., r/dermatomyositis), but the vast majority of mentions occurs in more general health related forums (e.g., r/AskDocs) or general autoimmunity related subreddits (e.g., r/Autoimmune).

```{r}
#| message: false
#| echo: false

library(knitr)
library(kableExtra)
subr <- df |>
  count(subreddit) |>
  rename(freq = n) |>
  mutate(percentage = (freq / sum(freq)) * 100)|>
  arrange(desc(freq)) |>
  mutate(cum_percentage = cumsum(percentage)) |>
  mutate(percentage = round(percentage, 1),
         cum_percentage = round(cum_percentage,1)) |>
  slice_head(n = 10) # Select top 10 subreddits

# Create table
subr |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

The table below shows the occurrences of myositis related words across all submissions.

```{r}
df$text <- paste(df$title, df$selftext, sep = " ")

# Assuming your data frame is called 'df' and the text column is called 'text'
text_col <- tolower(df$text)
text_col <- gsub("_", " ", text_col)

# Define a regular expression pattern to match words containing "myositis"
pattern <- "\\b\\w*myositis\\w*\\b"

# Find all matches in the text column
matches <- str_extract_all(text_col, pattern)

# Create a vector of all matched words
matched_words <- unlist(matches)

# Remove duplicates and count occurrences
word_counts <- table(matched_words)

# Convert the table to a data frame
word_counts_df <- as.data.frame(word_counts)
colnames(word_counts_df) <- c("Word", "Count")

# Sort the data frame by the count in descending order
word_counts_df <- word_counts_df[order(-word_counts_df$Count), ]

word_counts_df <- word_counts_df |>
  mutate(percentage = (Count / sum(Count)) * 100)|>
  arrange(desc(Count)) |>
  mutate(cum_percentage = cumsum(percentage)) |>
  mutate(percentage = round(percentage, 1),
         cum_percentage = round(cum_percentage,1)) |>
  slice_head(n = 20) # Select top 10 subreddits

# Print the result
word_counts_df |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r}
pronoun_pattern <- "\\b(i|me|my|she|he|her|his)\\b"

# Filter matching rows
df_pronouns <- df |>
  filter(str_detect(text, regex(pronoun_pattern, ignore_case = TRUE)))

df_no_pronouns <- df |>
  filter(!str_detect(text, regex(pronoun_pattern, ignore_case = TRUE)))

```
