---
title: "Myositis and Pain: Method & Results"
format:
  html: default
  pdf:
    documentclass: article
    geometry: a4paper
  docx: default
editor: visual
execute:
  echo: false
  warning: false
  message: false
editor_options:
  chunk_output_type: inline
always_allow_html: true
---

```{r}
#| label: setup-load
#| message: false

# ------------------------------------------------------------
# Project: Myositis & Pain (Reddit text + LLM labels)
# Purpose: Descriptive analyses + pain lexicon matching + semantic network
# Notes for sharing:
#   - This QMD assumes you have a processed dataset already (see cleaning script path below).
#   - It reads one Feather file from /LLM output/.
#   - It pulls the pain lexicon from a public Google Sheet (read-only).
# ------------------------------------------------------------

# Clean workspace (keep loaded functions)
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")

options(scipen = 999)  # avoid scientific notation in tables/prints

# ---- Libraries ----
# Tip: For GitHub sharing, it’s useful to keep all libraries in one place.
library(tidyverse)
library(tidytext)
library(SnowballC)
library(textstem)

library(purrr)
library(forcats)
library(viridis)

library(ggvenn)
library(RColorBrewer)
library(wordcloud)

library(widyr)
library(igraph)
library(ggraph)
library(tidygraph)
library(tidylo)

library(here)
library(arrow)
library(scales)
library(gt)
library(glue)

library(htmltools)
library(gtExtras)
library(ggthemes)

library(googlesheets4)
library(lubridate)   # floor_date() is used later
library(knitr)

df <- arrow::read_feather(here("LLM output", "majority_vote_gpt51_20251210_181138.feather"))
```

```{r}
#| label: descriptives-over-time
#| message: false

# ------------------------------------------------------------
# Descriptives over time
# 1) Number of posts per month
# 2) Subreddit composition over years (proportions)
# ------------------------------------------------------------

# ---- 1) Mentions per month ----
# Assumption: created_utc can be safely converted to Date (already Date or POSIXct/character).
df_summary <- df |>
  mutate(year_month = floor_date(as.Date(created_utc), "month")) |>
  group_by(year_month) |>
  summarise(count = n(), .groups = "drop")

ggplot(df_summary, aes(x = year_month, y = count)) +
  geom_point() +
  geom_jitter() +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Number of Mentions per Month",
    x = "Year",
    y = "Count"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 50, hjust = 1))

# ---- 2) Subreddits over time (composition) ----
top_n <- 7  # number of subreddits to show; everything else is lumped into "Other"

df_plot <- df |>
  mutate(
    year = as.integer(year),
    subreddit = fct_lump_n(subreddit, n = top_n, other_level = "Other")
  ) |>
  count(year, subreddit, name = "n_posts")

ggplot(df_plot, aes(x = year, y = n_posts, fill = subreddit)) +
  geom_area(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(
    x = "Year",
    y = "Proportion of posts",
    fill = "Subreddit",
    title = "Proportion of IIM-related posts per subreddit over time"
  ) +
  theme_minimal()

df_recent_years <- df |>
  filter(year == 2025) |>
  count(month, name = "n_posts")
mean(df_recent_years$n_posts)

perc_table(df, "year")
perc_table(df, "subreddit")

n_unique_authors <- length(unique(df$author))
n_unique_subreddits <- length(unique(df$subreddit))
n_posts <- nrow(df)
```

```{r}
#| label: preparing the pain lexicon, lemmatizing 
#| message: false

rm(df_plot, df_summary)

# ---- 1) Build pain lexicon ----
# Pain lexicon source:
# Chaturvedi, J., Mascio, A., Velupillai, S. U., & Roberts, A. (2021).
# Development of a Lexicon for Pain. Front Digit Health, 3, 778305.

gs4_deauth()  # ensures this works without requiring user auth (read-only public sheet)
sheet_url <- "https://docs.google.com/spreadsheets/d/1z-6619UBdvwWrB9Sz4b1rbjDzuslOGCpts2DNc0naCc/edit?gid=761302369#gid=761302369"
# Alternative for reproducibility: download the sheet and store it in /data/, then read locally.

pain_lexicon_original <- read_sheet(sheet_url) |>
  select(word = `keywords-clean`, num_words = `count of words`) |>
  mutate(
    word = str_squish(tolower(as.character(word))),
    num_words = as.integer(num_words)
  ) |>
filter(!word %in% c("p ain", "p ainful", "p ains"))

# Project-specific additions (kept explicit so readers can audit changes)
custom_pain_keywords <- c(
  # Phrases
  "extreme pain", "constant pain", "severe pain", "sharp pain", "achy pain", "terrible pain",
  "intense pain", "worst pain", "unusual pain", "horrible pain", "weird pain", "dull pain",
  "mild pain", "persistent pain", "significant pain", "unbearable pain", "excruciating pain",
  "searing pain", "intolerable pain", "piercing pain", "burning pain", "throbbing pain",
  "stabbing pain", "shooting pain", "debilitating pain", "electric shock", "pins and needles",
  "burning sensation", "feels bruised", "limited mobility", "morning stiffness",
  # Single Words
  "sore", "hurt", "hurting", "discomfort", "burning", "throbbing", "stabbing", "shooting",
  "inflammation", "inflamed", "stiffness", "tender", "spasm", "migraine", "flare", "flares",
  "flare-up", "flareup", "agony", "unbearable", "excruciating", "debilitating", "torture",
  "torment", "intolerable", "suffering", "searing", "piercing", "twinge", "neuropathy",
  "neuropathic", "tingling", "numbness", "zapping", "hypersensitive", "locking", "tightness",
  "tenderness", "allodynia", "hyperalgesia"
)

pain_lexicon_custom <- tibble(word = custom_pain_keywords) |>
  mutate(
    word = str_squish(tolower(word)),
    num_words = str_count(word, "\\s+") + 1L
  )

pain_lexicon_raw <- bind_rows(pain_lexicon_original, pain_lexicon_custom) |>
  filter(!is.na(word), word != "", !is.na(num_words)) |>
  distinct(word, .keep_all = TRUE)

rm(pain_lexicon_original, pain_lexicon_custom)
# lemmatization for pain words 

library(udpipe)
model <- udpipe_load_model("english-ewt-ud-2.5-191206.udpipe") 

raw_pain_words <- pain_lexicon_raw |>
  filter(num_words == 1) |> #we are not using the bigrams or trigrams
  mutate(word = str_squish(str_to_lower(word))) |>
  distinct() |>
  pull(word)

# Define Anchors with manual corrections
clean_myositis_terms <- function(v) {
  case_when(
    # --- Ache Grouping ---
    v %in% c("ach", "achiness", "achyness", "ache") ~ "ache",
    v == "earach"   ~ "earache",
    v == "headach"  ~ "headache",
    v == "neckach"  ~ "neckache",

    # --- Medical Terminology ---
    v %in% c("cephalodynia", "cephalodynias") ~ "cephalodynia",
    v == "dysesthesia" ~ "dysaesthesia",
    v == "neuropathic" ~ "neuropathy",
    v == "inflamed"    ~ "inflammation",
    v == "inflam"    ~ "inflammation",
    v == "inflammatory" ~ "inflammation",
    v == "cephalalgia" ~ "cephalodynia",
    v == "anti synthetase" ~ "antisynthetase",
    
    # --- Tailbone Grouping ---
    v %in% c("coccyalgia", "coccygalgia", "coccygodynia", 
             "coccyodynia", "coccyxdynia") ~ "coccydynia",

    # --- Flare Handling ---
    v %in% c("flareup", "flare-up", "flare", "up", "-") ~ "flare",

    # --- Pain Descriptors (Adjective Preference) ---
    v == c("hurting", "hurts")    ~ "hurt",
    v == "burn"      ~ "burning",
    v == "throbb"    ~ "throbbing",
    v == "stab"      ~ "stabbing",
    v == "shoot"     ~ "shooting",
    v == "excruciat" ~ "excruciating",
    v == "debilitat" ~ "debilitating",
    v == "sear"      ~ "searing",
    v == "lock"      ~ "locking",
    v == "soreness" ~ "sore",
    v == "weak" ~ "weakness",

    # --- Fallback ---
    TRUE ~ v
  )
}

pain_lexicon_lemmatized <- udpipe_annotate(model, x = raw_pain_words) |>
  as.data.frame() |>
  mutate(lemma = clean_myositis_terms(lemma))

# Tokenize from 'df$lemmatized_text'
tokens_normalized <- df |>
  select(doc_id, text = lemmatized_text) |>
  unnest_tokens(word, text)

tokens_normalized <- tokens_normalized |>
  mutate(word = clean_myositis_terms(word))

rm(model)
```

```{r}
#| label: pain-mentions (unigrams only)
#| message: false

#unigrams for pain terms (lemmatized text)
unigram_lexicon_lemmatized <- pain_lexicon_lemmatized |> 
  distinct(lemma) |>
  rename(word = lemma)

# ---- 2) Count lexicon matches in text (overall frequency) ----
unigram_counts_lemmatized <- df |>
  select(doc_id, text) |>
  unnest_tokens(word, text, token = "words") |>
  mutate(word = str_squish(tolower(word))) |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  count(word, sort = TRUE)

all_pain_matches <- tokens_normalized |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  select(doc_id, word)

# 2. Calculate Term Counts and Diversity (Unique Types) per post
pain_stats_per_post <- all_pain_matches |>
  group_by(doc_id) |>
  summarise(
    n_pain_tokens = n()    # Total count (density numerator)
  ) |>
  ungroup()


# 3. Join back to your main dataframe
df <- df |>
  left_join(pain_stats_per_post, by = "doc_id") |>
  mutate(
    n_pain_tokens = replace_na(n_pain_tokens, 0),
  )

pain_counts_per_post <- all_pain_matches |>
  count(doc_id, name = "n_pain_terms")

df <- df |>
  left_join(pain_counts_per_post, by = "doc_id") |>
  mutate(n_pain_terms = replace_na(n_pain_terms, 0))

df_pain <- df |>
  filter(n_pain_terms > 0)

# Basic descriptives (kept as objects for later reporting)
n_posts_with_pain       <- nrow(df_pain)
prop_posts_with_pain    <- nrow(df_pain) / nrow(df)
n_authors_with_pain     <- length(unique(df_pain$author))
prop_authors_with_pain  <- length(unique(df_pain$author)) / length(unique(df$author))
n_pain_terms            <- sum(df_pain$n_pain_terms)


```

number of posts with pain `r n_posts_with_pain` proportion of posts with pain `r prop_posts_with_pain` number of authors with pain `r n_authors_with_pain` proportion of authors with pain `r prop_authors_with_pain` number of pain terms `r sum(df_pain$n_pain_terms)`

```{r}
#| label: pain-descriptions
#| message: false

# ---- Pain-experience vs pain-location (for bigrams ending in " pain") ----
# Heuristic: classify "<descriptor> pain" as pain_experience; otherwise pain_location.
pain_descriptors <- c(
  "chronic","extreme","constant","severe","sharp","burning","bad","achy","terrible",
  "intense","excruciating","worst","aching","cramping","unusual","stabbing",
  "debilitating","horrible","phantom","weird","dull","mild","persistent",
  "significant","shooting","acute","moderate","radiating","squeezing",
  "unbearable","agonizing","deep","heavy","hot","intermittent","migratory",
  "miserable","searing","superficial","throbbing","tingling","transient",
  "much", "extremely", "agonizingly", "crawling", "disabling", "excruciatingly",
  "excruciating", "horribly", "incredibly", "more", "most", "less", "pure", "random",
  "severely", "seething", "strange", "super", "unrelenting"
)

pain_bigrams <- df |>
  select(doc_id, text) |>
  mutate(text = str_replace_all(text, "[’']", "")) |>
  mutate(
    bigram = str_extract_all(
      text,
      regex("\\b(\\w+)\\s+(pain|painful)\\b", ignore_case = TRUE)
    )
  ) |>
  unnest(bigram) |>
  separate(bigram, into = c("word1", "word2"), sep = "\\s+") |>
  mutate(
    word1 = str_to_lower(word1),
    word2 = str_to_lower(word2)
  ) |>
  filter(!is.na(word1), !is.na(word2)) |>
  select(doc_id, word1, word2)

#remove stopwords in word1 + check
removed_by_stopwords <- pain_bigrams |>
  semi_join(stop_words, by = c("word1" = "word")) |>
  distinct(word1) |>
  arrange(word1) 

keep_words <- c(
  "back", "face", "little", "less", "much", "more", 
  "most", "serious")

pain_bigrams <- pain_bigrams |>
  filter(
    !word1 %in% stop_words$word |
    word1 %in% keep_words
  ) |>
    count(word1, word2, sort = TRUE)

pain_bigrams <- pain_bigrams |>
  mutate(
    category = if_else(word1 %in% pain_descriptors, "pain_experience", "other")
  )

bigram_experience_counts <- pain_bigrams |>
  filter(category == "pain_experience") |>
  unite("bigram", word1, word2, sep = " ") |>
  select(bigram, n) |>
  arrange(desc(n))
```

```{r}
#| label: tables pain-terms and pain-descriptions
#| message: false

top_n_terms   <- 15
top_n_bigrams <- 15

# Top lists + relative/cumulative percentages (within panel) ----
top_terms <- unigram_counts_lemmatized |>
  slice_max(n, n = top_n_terms, with_ties = FALSE) |>
  arrange(desc(n)) |>
  mutate(
    rel = (n / sum(unigram_counts_lemmatized$n)) * 100,
    cum = cumsum(rel)
  ) |>
  transmute(
    term = word,
    term_n = n,
    term_rel = round(rel, 1),
    term_cum = round(cum, 1),
    row = row_number()
  )

top_bigrams <- bigram_experience_counts |>
  slice_max(n, n = top_n_bigrams, with_ties = FALSE) |>
  arrange(desc(n)) |>
  mutate(
    rel = (n / sum(bigram_experience_counts$n)) * 100,
    cum = cumsum(rel)
  ) |>
  transmute(
    bigram = bigram,
    bigram_n = n,
    bigram_rel = round(rel, 1),
    bigram_cum = round(cum, 1),
    row = row_number()
  )

top_side_by_side <- full_join(top_terms, top_bigrams, by = "row") |>
  arrange(row) |>
  select(-row)

# ---- 5) GT table ----
top_side_by_side_gt <- top_side_by_side |>
  gt() |>
  tab_header(
    title = "Top pain-related terms and pain-experience bigrams",
    subtitle = "Identified in myositis-related Reddit posts"
  ) |>
  cols_label(
    term = "Top pain term (unigram)",
    term_n = "n",
    term_rel = "Relative (%)",
    term_cum = "Cumulative (%)",
    bigram = "Top pain-experience bigram",
    bigram_n = "n",
    bigram_rel = "Relative (%)",
    bigram_cum = "Cumulative (%)"
  ) |>
  fmt_number(columns = c(term_n, bigram_n), decimals = 0) |>
  fmt_number(columns = c(term_rel, term_cum, bigram_rel, bigram_cum), decimals = 1) |>
  cols_align(align = "left", columns = c(term, bigram)) |>
  cols_align(align = "right", columns = c(term_n, term_rel, term_cum, bigram_n, bigram_rel, bigram_cum)) |>
  tab_source_note(
    source_note = md(
      "n indicates the total number of occurrences across all posts.<br>
       Relative (%) and Cumulative (%) are computed within unigrams and within pain-experience bigrams, respectively.<br>
       Bigrams are restricted to phrases ending in 'pain' and classified as pain-experience when the first token is a descriptor (e.g., severe, burning), otherwise as pain-location (e.g., neck pain)."
    )
  )

top_side_by_side_gt

```

```{r}
#| label: body-regions-in-pain-posts
#| message: false

# ------------------------------------------------------------
# Body regions mentioned in pain posts
#
# Important assumption:
# - This chunk uses df_pain$lemmatized_text (already prepared upstream).
#   If you share this on GitHub, document where lemmatized_text is created
#   (e.g., in your cleaning pipeline).
# ------------------------------------------------------------

library(pluralize)

body_parts_singular <- c(
  "head", "face", "forehead", "eye", "eyelid", "eyebrow", "eyelash", "ear", "nose",
  "nostril", "mouth", "lip", "tongue", "tooth", "gum", "chin", "cheek", "jaw",
  "scalp", "temple", "neck", "throat", "larynx", "adam's apple", "shoulder", "arm",
  "elbow", "forearm", "wrist", "hand", "finger", "thumb", "palm", "knuckle", "nail",
  "chest", "breast", "rib", "abdomen", "stomach", "navel", "waist", "back", "spine",
  "hip", "pelvis", "leg", "thigh", "knee", "calf", "shin", "ankle", "foot", "heel",
  "toe", "sole", "heart", "lung", "liver", "kidney", "brain", "bladder", "pancreas",
  "intestine", "colon", "appendix", "spleen", "gallbladder", "joint", "bone",
  "muscle", "ligament", "tendon", "skin", "hair", "blood vessel", "vein", "artery",
  "nerve", "penis", "vagina"
)

# 1. Create the expanded dictionary for matching
body_parts_expanded <- tibble(word = body_parts_singular) |>
  mutate(plural = pluralize(word)) |>
  pivot_longer(cols = everything(), values_to = "match_word") |>
  distinct(match_word)

# 2. Join, Normalize, and Count
body_part_counts <- df_pain |>
  unnest_tokens(output = word, input = lemmatized_text) |>
  # Match against the expanded list
  inner_join(body_parts_expanded, by = c("word" = "match_word")) |>
  # Normalize: Convert "eyes" back to "eye" so they group together
  mutate(word = singularize(word)) |> 
  count(word, sort = TRUE) |>
  mutate(
    relative_percentage = (n / sum(n)) * 100,
    cumulative_percentage = cumsum(relative_percentage),
    relative_percentage = round(relative_percentage, 1),
    cumulative_percentage = round(cumulative_percentage, 1)
  )

top_20_table_body_regions <- body_part_counts |>
  slice_max(order_by = n, n = 20) |>
  gt() |>
  tab_header(
    title = md("**Top 20 Body Regions mentioned in Pain Related Myositis posts**")
  ) |>
  cols_label(
    word = "Body Region",
    n = "n",
    relative_percentage = "Relative (%)",
    cumulative_percentage = "Cumulative (%)"
  ) |>
  fmt_number(columns = n, decimals = 0) |>
  tab_source_note(
    source_note = md(
      "n indicates the total number of occurrences across all posts with at least one pain term.<br>
       Relative (%) and Cumulative (%) indicate the proportion and cumulative proportion of all body-region mentions, respectively."
    )
  )

top_20_table_body_regions

# Posts that mention *any* body part (doc-level indicator)
posts_with_bodypart <- df_pain |>
  unnest_tokens(output = word, input = lemmatized_text) |>
  inner_join(tibble(word = body_parts_singular), by = "word") |>
  distinct(doc_id)

n_posts_with_bodypart <- nrow(posts_with_bodypart)
prop_posts_with_bodypart <- n_posts_with_bodypart / nrow(df_pain)
```

number of pain posts with body regions `r nrow(posts_with_bodypart)` number of pain posts with body regions `r n_posts_with_bodypart / nrow(df_pain)`

```{r}
#| label: pain-centrality-score
#| message: false

df_pain <- df_pain |> #remove old n_tokens column
  select(-n_tokens)

# 1) Post length in tokens
post_len <- tokens_normalized |>
  count(doc_id, name = "n_tokens")

# 2) Pain matches per post: occurrences + distinct terms
pain_per_post <- tokens_normalized |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  group_by(doc_id) |>
  summarise(
    n_pain_occurrences = n(),
    n_pain_distinct   = n_distinct(word),
    .groups = "drop"
  )

# 3) Join back to df_pain + compute metrics
df_pain <- df_pain |>
  left_join(post_len, by = "doc_id") |>
  left_join(pain_per_post, by = "doc_id") |>
  mutate(
    n_tokens           = coalesce(n_tokens, 0L),
    n_pain_occurrences = coalesce(n_pain_occurrences, 0L),
    n_pain_distinct    = coalesce(n_pain_distinct, 0L),
    pain_density       = n_pain_occurrences / pmax(n_tokens, 1),
    pain_centrality    = pain_density * log1p(n_pain_distinct)
  ) |>
  arrange(desc(pain_centrality))

```

• pain_density = pain-term occurrences / total tokens\
• pain_diversity = number of distinct pain terms used\
• pain_centrality = pain_density \* log1p(pain_diversity)

```{r}
#| label: prepare semantic network
#| message: false

# Define Stop Words (Visual clutter removal)
# Since the lemmatization already removed basic stopwords (the/and)
# this list focuses on words that are "content" but uninteresting (e.g., "doctor").

custom_stop_words <- tibble(
  word = c(
    "time","day","week","month","year","months","days", "hour",
    "lot", "bit", "post", "doctor", "doctors", "wait",
    "diagnosed", "diagnosis", "symptoms", "disease", "body",
    "people", "person", "issue", "issues", "result", "results",
    "start", "started", "feel", "feeling", "bad", "worse",
    "life", "happen", "normal", "live", "read", "symptom", "experience",
    "diagnose", "medical", "notice", "health", "syndrome", "low", "include",
    "deal", "level", "typical", "refuse", "leave"
  )
)

# We define broad disease terms to anchor the network
disease_pattern <- "(myositis|antisynthetase|myopath)"

top_pain_docs <- df_pain |>
  arrange(desc(pain_centrality)) |>
  slice_head(n = 500)

tokens_normalized_filtered <- tokens_normalized |>
  semi_join(top_pain_docs |> select(doc_id), by = "doc_id") |>
  anti_join(custom_stop_words, by = "word") |>
  filter(
    str_length(word) > 2,
    !str_detect(word, "^[0-9]+$")
  )

# 4. Document Frequencies
term_doc_freq <- tokens_normalized_filtered |>
  distinct(doc_id, word) |>
  count(word, name = "df")

# 5. Pairwise Co-occurrence
pairs <- tokens_normalized_filtered |>
  distinct(doc_id, word) |>
  widyr::pairwise_count(word, doc_id, sort = TRUE, upper = FALSE)

# 6. Jaccard & Filtering
pairs_jaccard <- pairs |>
  left_join(term_doc_freq, by = c("item1" = "word")) |>
  rename(df1 = df) |>
  left_join(term_doc_freq, by = c("item2" = "word")) |>
  rename(df2 = df) |>
  mutate(jaccard = n / (df1 + df2 - n))
```

```{r}
#| label: semantic-networks-pain
#| message: false

min_jaccard <- 0.14  
min_node_n  <- 13       
max_node_n  <- 350      
label_min_n <- 13
degree_filter <- 1

pain_terms <- unigram_lexicon_lemmatized$word

# Filter: Edges must involve a Pain Term OR a Disease Term
pain_edges <- pairs_jaccard |>
  filter(
    item1 %in% pain_terms | item2 %in% pain_terms |
    str_detect(item1, disease_pattern) | str_detect(item2, disease_pattern)
  ) |>
  filter(jaccard >= min_jaccard) |>
  transmute(from = item1, to = item2, weight = jaccard)

# 7. Nodes & Graph Build
pain_nodes <- tibble(name = unique(c(pain_edges$from, pain_edges$to))) |>
  left_join(term_doc_freq, by = c("name" = "word")) |>
  mutate(
    df = replace_na(df, 0L),
    is_pain_term = name %in% pain_terms
  ) |>
  filter(df >= min_node_n, df <= max_node_n)

pain_edges <- pain_edges |>
  filter(from %in% pain_nodes$name & to %in% pain_nodes$name)

library(tidygraph)
g <- tbl_graph(nodes = pain_nodes, edges = pain_edges, directed = FALSE) |>
  activate(nodes) |>
  mutate(deg = centrality_degree()) |>
  filter(deg >= degree_filter) 

# 8. Plot (Same as your original script)
library(grid)
set.seed(42)
p <- ggraph(g, layout = "stress") +
  geom_edge_link(aes(width = weight), alpha = 0.3) +
  geom_node_point(aes(color = df, shape = is_pain_term), size = 5) +
  geom_node_text(
    aes(label = name, filter = df >= label_min_n),
    repel = TRUE, size = 3
  ) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_color_viridis_c(option = "C", direction = -1) +
  scale_shape_manual(
    name = "Term type",
    values = c(`TRUE` = 17, `FALSE` = 16),
    labels = c(`TRUE` = "Pain term", `FALSE` = "Non-pain term")
  ) +
  guides(
    color = guide_colorbar(order = 1, barheight = unit(35, "mm")),
    shape = guide_legend(order = 2, override.aes = list(color = "#E69F00")),
    edge_width = guide_legend(order = 3)
  ) +
  labs(
    title = "Pain- and disease-anchored semantic network",
    subtitle = "Top 500 Myositis-related posts by combined pain centrality",
    caption = "Edges represent normalized co-occurrence of terms across posts.",
    edge_width = "Co-occurrence strength\n(normalized)", # Added \n to wrap long text
    color = "Posts with term"
  ) +
  theme_minimal(base_size = 12) +
  coord_cartesian(clip = "off") +
  theme(
    axis.text  = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),

    # --- FIX 1: Auto-place legend to the right ---
    legend.position = "right", 
    legend.box = "vertical",
    
    # --- FIX 2: Simplify margins ---
    # We no longer need huge right margin because 'legend.position="right"' 
    # automatically resizes the plot to make room.
    plot.margin = margin(10, 10, 10, 10),

    # Optional: Aligns title to the very left edge of image, not just the plot area
    plot.title.position = "plot", 
    
    plot.caption = element_text(size = 9, color = "grey40", hjust = 0)
  )

p2 <- ggraph(g, layout = "stress") +
  geom_edge_link(aes(width = weight), alpha = 0.3) +
  geom_node_point(aes(color = df, shape = is_pain_term), size = 5) +
  geom_node_text(
    aes(label = name, filter = df >= label_min_n),
    repel = TRUE, size = 3
  ) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_color_viridis_c(option = "C", direction = -1) +
  scale_shape_manual(
    name = "Term type",
    values = c(`TRUE` = 17, `FALSE` = 16),
    labels = c(`TRUE` = "Pain term", `FALSE` = "Non-pain term")
  ) +
  guides(
    color = guide_colorbar(order = 1, barheight = unit(35, "mm")),
    shape = guide_legend(order = 2, override.aes = list(color = "#E69F00")),
    edge_width = guide_legend(order = 3)
  ) +
  labs(
    edge_width = "Co-occurrence strength\n(normalized)", # Added \n to wrap long text
    color = "Posts with term"
  ) +
  theme_minimal(base_size = 12) +
  coord_cartesian(clip = "off") +
  theme(
    axis.text  = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),

    # --- FIX 1: Auto-place legend to the right ---
    legend.position = "right", 
    legend.box = "vertical",
    
    # --- FIX 2: Simplify margins ---
    # We no longer need huge right margin because 'legend.position="right"' 
    # automatically resizes the plot to make room.
    plot.margin = margin(10, 10, 10, 10),

    # Optional: Aligns title to the very left edge of image, not just the plot area
    plot.title.position = "plot", 
    
    plot.caption = element_text(size = 9, color = "grey40", hjust = 0)
  )

p2 
#ggsave(
#  filename = "Myositis_Semantic_Network_HighRes.png",
#  plot = p2,
#  width = 12,       # Width in inches
#  height = 8,       # Height in inches
#  dpi = 300,        # High resolution for publication
#  bg = "white"      # Ensures transparent backgrounds don't turn black
#)

cat("Nodes:", gorder(g), "\n")
cat("Edges:", gsize(g), "\n")
cat("Min/median/max df:\n")
print(summary(as_tibble(g, active="nodes")$df))
cat("Degree summary:\n")
print(summary(as_tibble(g, active="nodes")$deg))
```

```{r}

# 1. Calculate Centrality Measures from your existing graph object 'g'
centrality_stats <- g |>
  activate(nodes) |>
  mutate(
    betweenness = centrality_betweenness(), # Measure of being a "bridge"
    degree = centrality_degree()          # Measure of raw connection count
  ) |>
  as_tibble() |> # Extract node data to a regular dataframe
  arrange(desc(betweenness)) |> # Sort by most "bridging" terms first
  slice_head(n = 10) 

# 2. Format for a clean manuscript table
centrality_table <- centrality_stats |>
  mutate(
    Type = if_else(is_pain_term, "Pain Term", "General/Clinical Term"),
    betweenness = round(betweenness, 1)
  ) |>
  select(
    Term = name, 
    Type, 
    `Betweenness Centrality` = betweenness, 
    `Degree (Connections)` = degree
  )

# 3. Display the table
kable(centrality_table, 
      caption = "Top 10 Most Central Terms in the Myositis Pain Network")


```

```{r}
# ---- Community detection + composition analysis  ----


library(igraph)
gi <- as.igraph(g)

# 1) Community detection (Louvain; robust default for weighted, undirected graphs)
set.seed(123)
comm <- cluster_louvain(g, weights = E(g)$weight)
V(g)$community <- membership(comm)

# Key descriptives (reportable)
n_communities <- length(unique(V(g)$community))
Q_modularity  <- modularity(comm)

cat("Communities:", n_communities, "\n")
cat("Modularity (Louvain):", round(Q_modularity, 3), "\n")

# 2) Community composition summary table
community_summary <- tibble(
  term      = V(g)$name,
  community = V(g)$community,
  is_pain   = V(g)$is_pain_term,
  df        = V(g)$df
) |>
  group_by(community) |>
  summarise(
    n_terms        = n(),
    pain_terms     = sum(is_pain, na.rm = TRUE),
    non_pain_terms = sum(!is_pain, na.rm = TRUE),
    pain_share     = mean(is_pain, na.rm = TRUE),
    mean_df        = mean(df, na.rm = TRUE),
    top_terms      = paste(term[order(df, decreasing = TRUE)][1:min(7, n())],
                           collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(desc(n_terms))


# 3) Statistical test: do communities differ in pain vs non-pain composition?
contingency <- table(
  Community = V(g)$community,
  PainTerm  = V(g)$is_pain_term
)

# Use Fisher if any expected counts are small; otherwise chi-square is fine.
chi <- suppressWarnings(chisq.test(contingency))
if (any(chi$expected < 5)) {
  test <- fisher.test(contingency)
  test_name <- "Fisher's exact test"
} else {
  test <- chi
  test_name <- "Chi-square test"
}

cat("\nComposition test:", test_name, "\n")
print(test)

# 4) Optional: plot network colored by community (for interpretation/appendix)
p_comm <- ggraph(g, layout = "stress") +
  geom_edge_link(aes(width = weight), alpha = 0.2) +
  geom_node_point(aes(color = factor(community), shape = is_pain_term), size = 4) +
  scale_edge_width(range = c(0.2, 2.5)) +
  labs(
    title = "Pain- and disease-anchored semantic network",
    subtitle = paste0(
      "Communities: ", n_communities,
      " | Modularity: ", round(Q_modularity, 3)
    ),
    edge_width = "Co-occurrence strength (normalized)",
    color = "Semantic community",
    shape = "Term type"
  ) +
  scale_shape_manual(
    name = "Term type",
    values = c(`TRUE` = 17, `FALSE` = 16),
    labels = c(`TRUE` = "Pain term", `FALSE` = "Non-pain term")
  ) +
  theme_void()

print(p_comm)
print(community_summary)

```

```{r}
#additional test: Are pain/disease terms distributed across communities more than expected by chance?

set.seed(123)
n_perm <- 1000
pain_share_null <- numeric(n_perm)

for (i in seq_len(n_perm)) {
  shuffled <- sample(V(g)$is_pain_term)
  pain_share_null[i] <- tapply(
    shuffled,
    V(g)$community,
    mean
  ) |> var(na.rm = TRUE)
}

observed_var <- tapply(
  V(g)$is_pain_term,
  V(g)$community,
  mean
) |> var(na.rm = TRUE)

mean(pain_share_null >= observed_var)

table(V(g)$community, V(g)$is_pain_term)
```
