---
title: "Myositis and Pain: Analysis Pipeline"
subtitle: "Descriptive analyses, pain lexicon matching, and semantic network analysis"
author: "Peter Kerkhof"
date: last-modified
format:
  html: default
  pdf:
    documentclass: article
    geometry: a4paper
  docx: default
editor: visual
execute:
  echo: false
  warning: false
  message: false
editor_options:
  chunk_output_type: inline
always_allow_html: true
---

## Overview

This script analyzes myositis-related Reddit posts to understand pain experiences. The analysis includes:

1.  **Descriptive statistics** - Posts over time, subreddit composition
2.  **Pain lexicon matching** - Using a validated pain lexicon plus custom terms
3.  **Pain expression analysis** - Unigrams, bigrams, body regions
4.  **Semantic network analysis** - Co-occurrence patterns and community detection

## 1. Setup and Data Loading

```{r}
#| label: setup-environment
#| message: false

# -----------------------------------------------------------------------------
# ENVIRONMENT SETUP
# -----------------------------------------------------------------------------
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")

options(scipen = 999)  # Avoid scientific notation in tables/outputs
```

```{r}
#| label: load-libraries
#| message: false

# -----------------------------------------------------------------------------
# LOAD REQUIRED LIBRARIES
# -----------------------------------------------------------------------------

# Data manipulation
library(tidyverse)
library(purrr)
library(forcats)
library(lubridate)

# Text analysis
library(tidytext)
library(SnowballC)
library(textstem)
library(widyr)
library(udpipe)
library(pluralize)

# Network analysis
library(igraph)
library(ggraph)
library(tidygraph)
library(tidylo)

# Visualization
library(viridis)
library(ggvenn)
library(RColorBrewer)
library(wordcloud)
library(ggthemes)
library(grid)

# Tables and reporting
library(gt)
library(gtExtras)
library(knitr)
library(glue)
library(scales)
library(htmltools)

# File I/O
library(here)
library(arrow)
library(googlesheets4)

# Load project configuration (DATA_PATH)
source(here("config.R"))
```

```{r}
#| label: load-data
#| message: false

# -----------------------------------------------------------------------------
# LOAD CLASSIFIED DATA
# -----------------------------------------------------------------------------
# This dataset contains posts classified as "Yes" (author refers to own myositis)
# by GPT-5.1 majority vote, after near-duplicate removal.

df <- arrow::read_feather(
  file.path(DATA_PATH, "LLM output", "majority_vote_gpt51_20251210_181138.feather")
)

cat("Loaded", scales::comma(nrow(df)), "posts\n")
cat("Date range:", min(df$date), "to", max(df$date), "\n")
cat("Unique authors:", scales::comma(length(unique(df$author))), "\n")
cat("Unique subreddits:", scales::comma(length(unique(df$subreddit))), "\n")
```

```{r}
#| label: helper-functions
#| message: false

# -----------------------------------------------------------------------------
# HELPER FUNCTIONS
# -----------------------------------------------------------------------------

#' Create a percentage table for a categorical variable
#'
#' @param data Data frame
#' @param var_name Column name (string)
#' @return GT table with counts and percentages

perc_table <- function(data, var_name) {
  data |>
    count(.data[[var_name]], name = "n") |>
    mutate(
      percentage = round(n / sum(n) * 100, 1),
      cum_percentage = round(cumsum(percentage), 1)
    ) |>
    arrange(desc(n)) |>
    gt() |>
    fmt_number(columns = n, decimals = 0) |>
    tab_header(title = paste("Distribution of", var_name))
}
```

## 2. Descriptive Statistics Over Time

```{r}
#| label: posts-per-month
#| message: false

# -----------------------------------------------------------------------------
# POSTS PER MONTH
# -----------------------------------------------------------------------------

df_summary <- df |>
  mutate(year_month = floor_date(as.Date(created_utc), "month")) |>
  group_by(year_month) |>
  summarise(count = n(), .groups = "drop")

ggplot(df_summary, aes(x = year_month, y = count)) +
  geom_point() +
  geom_line(alpha = 0.3) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Myositis-Related Posts per Month",
    subtitle = "Posts where author refers to their own myositis diagnosis",
    x = "Year",
    y = "Number of Posts"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: subreddit-composition
#| message: false

# -----------------------------------------------------------------------------
# SUBREDDIT COMPOSITION OVER TIME
# -----------------------------------------------------------------------------

top_n_subreddits <- 7  # Show top N; others lumped into "Other"

df_plot <- df |>
  mutate(
    year = as.integer(year),
    subreddit = fct_lump_n(subreddit, n = top_n_subreddits, other_level = "Other")
  ) |>
  count(year, subreddit, name = "n_posts")

ggplot(df_plot, aes(x = year, y = n_posts, fill = subreddit)) +
  geom_area(position = "fill") +
  scale_y_continuous(labels = percent) +
  labs(
    x = "Year",
    y = "Proportion of Posts",
    fill = "Subreddit",
    title = "Subreddit Composition of Myositis Posts Over Time"
  ) +
  theme_minimal()
```

```{r}
#| label: descriptive-stats
#| message: false

# -----------------------------------------------------------------------------
# SUMMARY STATISTICS
# -----------------------------------------------------------------------------

# Recent activity (2025)
df_recent_years <- df |>
  filter(year == 2025) |>
  count(month, name = "n_posts")

mean_posts_per_month_2025 <- mean(df_recent_years$n_posts)

# Store key counts for reporting
n_posts <- nrow(df)
n_unique_authors <- length(unique(df$author))
n_unique_subreddits <- length(unique(df$subreddit))

# Display tables
perc_table(df, "year")
perc_table(df, "subreddit")
```

**Dataset Summary:**

-   Total posts: `r scales::comma(n_posts)`
-   Unique authors: `r scales::comma(n_unique_authors)`
-   Unique subreddits: `r scales::comma(n_unique_subreddits)`
-   Average posts per month (2025): `r round(mean_posts_per_month_2025, 1)`

## 3. Pain Lexicon Preparation

```{r}
#| label: load-pain-lexicon
#| message: false

# -----------------------------------------------------------------------------
# LOAD PAIN LEXICON
# -----------------------------------------------------------------------------
# Source: Chaturvedi, J., Mascio, A., Velupillai, S. U., & Roberts, A. (2021).
# Development of a Lexicon for Pain. Front Digit Health, 3, 778305.

# Clean up temporary objects
rm(df_plot, df_summary)

# Read from public Google Sheet (read-only, no auth required)
gs4_deauth()
sheet_url <- "https://docs.google.com/spreadsheets/d/1z-6619UBdvwWrB9Sz4b1rbjDzuslOGCpts2DNc0naCc/edit?gid=761302369#gid=761302369"

pain_lexicon_original <- read_sheet(sheet_url) |>
  select(word = `keywords-clean`, num_words = `count of words`) |>
  mutate(
    word = str_squish(tolower(as.character(word))),
    num_words = as.integer(num_words)
  ) |>
  filter(!word %in% c("p ain", "p ainful", "p ains")) |>  # Remove OCR errors
  mutate(source = "Chaturvedi et al")

#writexl::write_xlsx(pain_lexicon_original, file.path(DATA_PATH, "pain_lexicons/pain_lexicon_original.xlsx"))
```

```{r}
#| label: custom-pain-terms
#| message: false

# -----------------------------------------------------------------------------
# ADD CUSTOM PAIN TERMS
# -----------------------------------------------------------------------------
# Project-specific additions to capture myositis-relevant pain expressions

custom_pain_keywords <- c(
  # Pain intensity phrases
  "extreme pain", "constant pain", "severe pain", "sharp pain", "achy pain",
  "terrible pain", "intense pain", "worst pain", "unusual pain", "horrible pain",
  "weird pain", "dull pain", "mild pain", "persistent pain", "significant pain",
  "unbearable pain", "excruciating pain", "searing pain", "intolerable pain",
  "piercing pain", "burning pain", "throbbing pain", "stabbing pain",
  "shooting pain", "debilitating pain",

  # Sensation descriptors
  "electric shock", "pins and needles", "burning sensation", "feels bruised",
  "limited mobility", "morning stiffness",

  # Single-word pain terms
  "sore", "hurt", "hurting", "discomfort", "burning", "throbbing", "stabbing",
  "shooting", "inflammation", "inflamed", "stiffness", "tender", "spasm",
  "migraine", "flare", "flares", "flare-up", "flareup", "agony", "unbearable",
  "excruciating", "debilitating", "torture", "torment", "intolerable",
  "suffering", "searing", "piercing", "twinge", "neuropathy", "neuropathic",
  "tingling", "numbness", "zapping", "hypersensitive", "locking", "tightness",
  "tenderness", "allodynia", "hyperalgesia"
)

pain_lexicon_custom <- tibble(word = custom_pain_keywords) |>
  mutate(
    word = str_squish(tolower(word)),
    num_words = str_count(word, "\\s+") + 1L,
    source = "Custom")

#writexl::write_xlsx(pain_lexicon_custom, file.path(DATA_PATH, "pain_lexicons/pain_lexicon_custom.xlsx"))

# Combine original and custom lexicons
pain_lexicon_raw <- bind_rows(pain_lexicon_original, pain_lexicon_custom) |>
  filter(!is.na(word), word != "", !is.na(num_words)) |>
  distinct(word, .keep_all = TRUE)

#writexl::write_xlsx(pain_lexicon_raw, file.path(DATA_PATH, "pain_lexicons/pain_lexicon_raw.xlsx"))

rm(pain_lexicon_original, pain_lexicon_custom)

cat("Pain lexicon size:", nrow(pain_lexicon_raw), "terms\n")
cat("  Unigrams:", sum(pain_lexicon_raw$num_words == 1), "\n")
cat("  Bigrams+:", sum(pain_lexicon_raw$num_words > 1), "\n")
```

```{r}
#| label: lemmatize-pain-lexicon
#| message: false

# -----------------------------------------------------------------------------
# LEMMATIZE PAIN LEXICON
# -----------------------------------------------------------------------------
# Use udpipe to normalize pain terms to their lemma forms

model <- udpipe_load_model("english-ewt-ud-2.5-191206.udpipe")

# Extract unigrams only (bigrams handled separately)
raw_pain_words <- pain_lexicon_raw |>
  filter(num_words == 1) |>
  mutate(word = str_squish(str_to_lower(word))) |>
  distinct() |>
  pull(word)
```

```{r}
#| label: define-term-normalization
#| message: false

# -----------------------------------------------------------------------------
# DOMAIN-SPECIFIC TERM NORMALIZATION
# -----------------------------------------------------------------------------
# Manual corrections for lemmatization errors and domain-specific groupings

clean_myositis_terms <- function(v) {
  case_when(
    # Ache grouping
    v %in% c("ach", "achiness", "achyness", "ache") ~ "ache",
    v == "earach"   ~ "earache",
    v == "headach"  ~ "headache",
    v == "neckach"  ~ "neckache",

    # Medical terminology normalization
    v %in% c("cephalodynia", "cephalodynias") ~ "cephalodynia",
    v == "dysesthesia" ~ "dysaesthesia",
    v == "neuropathic" ~ "neuropathy",
    v %in% c("inflamed", "inflam", "inflammatory") ~ "inflammation",
    v == "cephalalgia" ~ "cephalodynia",
    v == "anti synthetase" ~ "antisynthetase",

    # Tailbone-related terms
    v %in% c("coccyalgia", "coccygalgia", "coccygodynia",
             "coccyodynia", "coccyxdynia") ~ "coccydynia",

    # Flare handling
    v %in% c("flareup", "flare-up", "flare", "up", "-") ~ "flare",

    # Pain descriptors (prefer adjective/present participle forms)
    v %in% c("hurting", "hurts") ~ "hurt",
    v == "burn"      ~ "burning",
    v == "throbb"    ~ "throbbing",
    v == "stab"      ~ "stabbing",
    v == "shoot"     ~ "shooting",
    v == "excruciat" ~ "excruciating",
    v == "debilitat" ~ "debilitating",
    v == "sear"      ~ "searing",
    v == "lock"      ~ "locking",
    v == "soreness"  ~ "sore",
    v == "weak"      ~ "weakness",

    # Default: keep as-is
    TRUE ~ v
  )
}

# Apply lemmatization and normalization to pain lexicon
pain_lexicon_lemmatized <- udpipe_annotate(model, x = raw_pain_words) |>
  as.data.frame() |>
  mutate(lemma = clean_myositis_terms(lemma))

# Tokenize post text and apply same normalization
tokens_normalized <- df |>
  select(doc_id, text = lemmatized_text) |>
  unnest_tokens(word, text) |>
  mutate(word = clean_myositis_terms(word))

rm(model)
```

## 4. Pain Mentions Analysis

```{r}
#| label: pain-unigram-counts
#| message: false

# -----------------------------------------------------------------------------
# COUNT PAIN TERM OCCURRENCES
# -----------------------------------------------------------------------------

# Create lookup table of lemmatized pain terms
unigram_lexicon_lemmatized <- pain_lexicon_lemmatized |>
  distinct(lemma) |>
  rename(word = lemma)

#writexl::write_xlsx(unigram_lexicon_lemmatized, file.path(DATA_PATH, "pain_lexicons/unigram_pain_terms_lemmatized.xlsx"))

# Count pain term occurrences in original text
unigram_counts_lemmatized <- df |>
  select(doc_id, text) |>
  unnest_tokens(word, text, token = "words") |>
  mutate(word = str_squish(tolower(word))) |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  count(word, sort = TRUE)

# Match pain terms in normalized tokens
all_pain_matches <- tokens_normalized |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  select(doc_id, word)
```

```{r}
#| label: pain-per-post
#| message: false

# -----------------------------------------------------------------------------
# CALCULATE PAIN STATISTICS PER POST
# -----------------------------------------------------------------------------

# Count pain tokens per post
pain_stats_per_post <- all_pain_matches |>
  group_by(doc_id) |>
  summarise(n_pain_tokens = n(), .groups = "drop")

# Join to main dataframe
df <- df |>
  left_join(pain_stats_per_post, by = "doc_id") |>
  mutate(n_pain_tokens = replace_na(n_pain_tokens, 0))

# Also count as n_pain_terms (for consistency)
pain_counts_per_post <- all_pain_matches |>
  count(doc_id, name = "n_pain_terms")

df <- df |>
  left_join(pain_counts_per_post, by = "doc_id") |>
  mutate(n_pain_terms = replace_na(n_pain_terms, 0))

# Create subset of posts with at least one pain term
df_pain <- df |>
  filter(n_pain_terms > 0)
```

```{r}
#| label: pain-descriptives
#| message: false

# -----------------------------------------------------------------------------
# PAIN DESCRIPTIVE STATISTICS
# -----------------------------------------------------------------------------

n_posts_with_pain      <- nrow(df_pain)
prop_posts_with_pain   <- nrow(df_pain) / nrow(df)
n_authors_with_pain    <- length(unique(df_pain$author))
prop_authors_with_pain <- length(unique(df_pain$author)) / length(unique(df$author))
n_pain_terms_total     <- sum(df_pain$n_pain_terms)
```

**Pain Mention Summary:**

-   Posts with pain terms: `r scales::comma(n_posts_with_pain)` (`r round(prop_posts_with_pain * 100, 1)`%)
-   Authors mentioning pain: `r scales::comma(n_authors_with_pain)` (`r round(prop_authors_with_pain * 100, 1)`%)
-   Total pain term occurrences: `r scales::comma(n_pain_terms_total)`

## 5. Pain Description Analysis

```{r}
#| label: pain-bigrams
#| message: false

# -----------------------------------------------------------------------------
# EXTRACT PAIN BIGRAMS
# -----------------------------------------------------------------------------
# Identify phrases like "severe pain", "chronic pain", etc.

# Define pain descriptor words (for classifying bigrams)
pain_descriptors <- c(
  "chronic", "extreme", "constant", "severe", "sharp", "burning", "bad", "achy",
  "terrible", "intense", "excruciating", "worst", "aching", "cramping", "unusual",
  "stabbing", "debilitating", "horrible", "phantom", "weird", "dull", "mild",
  "persistent", "significant", "shooting", "acute", "moderate", "radiating",
  "squeezing", "unbearable", "agonizing", "deep", "heavy", "hot", "intermittent",
  "migratory", "miserable", "searing", "superficial", "throbbing", "tingling",
  "transient", "much", "extremely", "agonizingly", "crawling", "disabling",
  "excruciatingly", "horribly", "incredibly", "more", "most", "less", "pure",
  "random", "severely", "seething", "strange", "super", "unrelenting"
)

# Extract bigrams ending in "pain" or "painful"
pain_bigrams <- df |>
  select(doc_id, text) |>
  mutate(text = str_replace_all(text, "['']", "")) |>
  mutate(
    bigram = str_extract_all(
      text,
      regex("\\b(\\w+)\\s+(pain|painful)\\b", ignore_case = TRUE)
    )
  ) |>
  unnest(bigram) |>
  separate(bigram, into = c("word1", "word2"), sep = "\\s+") |>
  mutate(
    word1 = str_to_lower(word1),
    word2 = str_to_lower(word2)
  ) |>
  filter(!is.na(word1), !is.na(word2)) |>
  select(doc_id, word1, word2)

# Remove stopwords but keep meaningful terms
keep_words <- c("back", "face", "little", "less", "much", "more", "most", "serious")

pain_bigrams <- pain_bigrams |>
  filter(!word1 %in% stop_words$word | word1 %in% keep_words) |>
  count(word1, word2, sort = TRUE) |>
  mutate(
    category = if_else(word1 %in% pain_descriptors, "pain_experience", "other")
  )

# Extract pain experience bigrams
bigram_experience_counts <- pain_bigrams |>
  filter(category == "pain_experience") |>
  unite("bigram", word1, word2, sep = " ") |>
  select(bigram, n) |>
  arrange(desc(n))
```

```{r}
#| label: pain-tables
#| message: false

# -----------------------------------------------------------------------------
# TOP PAIN TERMS AND BIGRAMS TABLE
# -----------------------------------------------------------------------------

top_n_terms   <- 15
top_n_bigrams <- 15

# Top unigrams with percentages
top_terms <- unigram_counts_lemmatized |>
  slice_max(n, n = top_n_terms, with_ties = FALSE) |>
  arrange(desc(n)) |>
  mutate(
    rel = (n / sum(unigram_counts_lemmatized$n)) * 100,
    cum = cumsum(rel)
  ) |>
  transmute(
    term = word,
    term_n = n,
    term_rel = round(rel, 1),
    term_cum = round(cum, 1),
    row = row_number()
  )

# Top bigrams with percentages
top_bigrams <- bigram_experience_counts |>
  slice_max(n, n = top_n_bigrams, with_ties = FALSE) |>
  arrange(desc(n)) |>
  mutate(
    rel = (n / sum(bigram_experience_counts$n)) * 100,
    cum = cumsum(rel)
  ) |>
  transmute(
    bigram = bigram,
    bigram_n = n,
    bigram_rel = round(rel, 1),
    bigram_cum = round(cum, 1),
    row = row_number()
  )

# Combine into side-by-side table
top_side_by_side <- full_join(top_terms, top_bigrams, by = "row") |>
  arrange(row) |>
  select(-row)

# Create GT table
top_side_by_side |>
  gt() |>
  tab_header(
    title = "Top Pain-Related Terms and Pain-Experience Bigrams",
    subtitle = "Identified in myositis-related Reddit posts"
  ) |>
  cols_label(
    term = "Pain Term (Unigram)",
    term_n = "n",
    term_rel = "Rel. (%)",
    term_cum = "Cum. (%)",
    bigram = "Pain-Experience Bigram",
    bigram_n = "n",
    bigram_rel = "Rel. (%)",
    bigram_cum = "Cum. (%)"
  ) |>
  fmt_number(columns = c(term_n, bigram_n), decimals = 0) |>
  fmt_number(columns = c(term_rel, term_cum, bigram_rel, bigram_cum), decimals = 1) |>
  cols_align(align = "left", columns = c(term, bigram)) |>
  cols_align(align = "right", columns = c(term_n, term_rel, term_cum, bigram_n, bigram_rel, bigram_cum)) |>
  tab_source_note(
    source_note = md(
      "n = total occurrences across all posts. Percentages computed within category.<br>
       Bigrams restricted to phrases ending in 'pain'; classified as pain-experience when first word is a descriptor (e.g., severe, burning)."
    )
  )
```

## 6. Body Regions in Pain Posts

```{r}
#| label: body-regions
#| message: false

# -----------------------------------------------------------------------------
# BODY REGION MENTIONS IN PAIN POSTS
# -----------------------------------------------------------------------------

# Define body parts vocabulary
body_parts_singular <- c(
  "head", "face", "forehead", "eye", "eyelid", "eyebrow", "eyelash", "ear",
  "nose", "nostril", "mouth", "lip", "tongue", "tooth", "gum", "chin", "cheek",
  "jaw", "scalp", "temple", "neck", "throat", "larynx", "adam's apple",
  "shoulder", "arm", "elbow", "forearm", "wrist", "hand", "finger", "thumb",
  "palm", "knuckle", "nail", "chest", "breast", "rib", "abdomen", "stomach",
  "navel", "waist", "back", "spine", "hip", "pelvis", "leg", "thigh", "knee",
  "calf", "shin", "ankle", "foot", "heel", "toe", "sole", "heart", "lung",
  "liver", "kidney", "brain", "bladder", "pancreas", "intestine", "colon",
  "appendix", "spleen", "gallbladder", "joint", "bone", "muscle", "ligament",
  "tendon", "skin", "hair", "blood vessel", "vein", "artery", "nerve",
  "penis", "vagina"
)

# Expand to include plurals
body_parts_expanded <- tibble(word = body_parts_singular) |>
  mutate(plural = pluralize(word)) |>
  pivot_longer(cols = everything(), values_to = "match_word") |>
  distinct(match_word)

# Count body part mentions in pain posts
body_part_counts <- df_pain |>
  unnest_tokens(output = word, input = lemmatized_text) |>
  inner_join(body_parts_expanded, by = c("word" = "match_word")) |>
  mutate(word = singularize(word)) |>  # Normalize to singular
  count(word, sort = TRUE) |>
  mutate(
    relative_percentage = round((n / sum(n)) * 100, 1),
    cumulative_percentage = round(cumsum(relative_percentage), 1)
  )

# Display top 20
body_part_counts |>
  slice_max(order_by = n, n = 20) |>
  gt() |>
  tab_header(
    title = md("**Top 20 Body Regions in Pain-Related Myositis Posts**")
  ) |>
  cols_label(
    word = "Body Region",
    n = "n",
    relative_percentage = "Rel. (%)",
    cumulative_percentage = "Cum. (%)"
  ) |>
  fmt_number(columns = n, decimals = 0) |>
  tab_source_note(
    source_note = md(
      "n = total occurrences in posts with at least one pain term.<br>
       Percentages indicate proportion of all body-region mentions."
    )
  )

library(writexl)

body_part_counts |>
  slice_max(order_by = n, n = 20) |>
  write_xlsx(file.path(DATA_PATH, "data/top20_body_regions_original.xlsx"))


```

```{r}
#| label: body-region-stats
#| message: false

# Count posts mentioning any body part
posts_with_bodypart <- df_pain |>
  unnest_tokens(output = word, input = lemmatized_text) |>
  inner_join(tibble(word = body_parts_singular), by = "word") |>
  distinct(doc_id)

n_posts_with_bodypart <- nrow(posts_with_bodypart)
prop_posts_with_bodypart <- n_posts_with_bodypart / nrow(df_pain)
```

**Body Region Summary:**

-   Pain posts mentioning body regions: `r scales::comma(n_posts_with_bodypart)` (`r round(prop_posts_with_bodypart * 100, 1)`%)

## 7. Pain Centrality Score

```{r}
#| label: pain-centrality
#| message: false

# -----------------------------------------------------------------------------
# CALCULATE PAIN CENTRALITY SCORE
# -----------------------------------------------------------------------------
# Combines pain density (frequency) with pain diversity (unique terms)
# Formula: pain_centrality = pain_density * log1p(pain_diversity)

# Remove old n_tokens if exists
if ("n_tokens" %in% names(df_pain)) {
  df_pain <- df_pain |> select(-n_tokens)
}

# Calculate post length in tokens
post_len <- tokens_normalized |>
  count(doc_id, name = "n_tokens")

# Calculate pain matches per post
pain_per_post <- tokens_normalized |>
  inner_join(unigram_lexicon_lemmatized, by = "word") |>
  group_by(doc_id) |>
  summarise(
    n_pain_occurrences = n(),
    n_pain_distinct = n_distinct(word),
    .groups = "drop"
  )

# Join and compute metrics
df_pain <- df_pain |>
  left_join(post_len, by = "doc_id") |>
  left_join(pain_per_post, by = "doc_id") |>
  mutate(
    n_tokens = coalesce(n_tokens, 0L),
    n_pain_occurrences = coalesce(n_pain_occurrences, 0L),
    n_pain_distinct = coalesce(n_pain_distinct, 0L),
    # Pain density: proportion of tokens that are pain terms
    pain_density = n_pain_occurrences / pmax(n_tokens, 1),
    # Pain centrality: density weighted by diversity
    pain_centrality = pain_density * log1p(n_pain_distinct)
  ) |>
  arrange(desc(pain_centrality))
```

**Pain Centrality Metrics:**

-   **pain_density** = pain-term occurrences / total tokens
-   **pain_diversity** = number of distinct pain terms used
-   **pain_centrality** = pain_density \* log1p(pain_diversity)

## 8. Semantic Network Analysis

```{r}
#| label: prepare-network
#| message: false

# -----------------------------------------------------------------------------
# PREPARE SEMANTIC NETWORK
# -----------------------------------------------------------------------------

# Define stop words for network visualization
# These are content words that are uninteresting for pain analysis
custom_stop_words <- tibble(
  word = c(
    "time", "day", "week", "month", "year", "months", "days", "hour",
    "lot", "bit", "post", "doctor", "doctors", "wait",
    "diagnosed", "diagnosis", "symptoms", "disease", "body",
    "people", "person", "issue", "issues", "result", "results",
    "start", "started", "feel", "feeling", "bad", "worse",
    "life", "happen", "normal", "live", "read", "symptom", "experience",
    "diagnose", "medical", "notice", "health", "syndrome", "low", "include",
    "deal", "level", "typical", "refuse", "leave"
  )
)

# Pattern for anchoring disease terms
disease_pattern <- "(myositis|antisynthetase|myopath)"

# Select top 500 posts by pain centrality for network
top_pain_docs <- df_pain |>
  arrange(desc(pain_centrality)) |>
  slice_head(n = 500)

# Filter tokens for network analysis
tokens_normalized_filtered <- tokens_normalized |>
  semi_join(top_pain_docs |> select(doc_id), by = "doc_id") |>
  anti_join(custom_stop_words, by = "word") |>
  filter(
    str_length(word) > 2,
    !str_detect(word, "^[0-9]+$")
  )

# Calculate document frequencies
term_doc_freq <- tokens_normalized_filtered |>
  distinct(doc_id, word) |>
  count(word, name = "df")

# Calculate pairwise co-occurrence
pairs <- tokens_normalized_filtered |>
  distinct(doc_id, word) |>
  widyr::pairwise_count(word, doc_id, sort = TRUE, upper = FALSE)

# Calculate Jaccard similarity
pairs_jaccard <- pairs |>
  left_join(term_doc_freq, by = c("item1" = "word")) |>
  rename(df1 = df) |>
  left_join(term_doc_freq, by = c("item2" = "word")) |>
  rename(df2 = df) |>
  mutate(jaccard = n / (df1 + df2 - n))
```

```{r}
#| label: build-network
#| message: false

# -----------------------------------------------------------------------------
# BUILD AND VISUALIZE SEMANTIC NETWORK
# -----------------------------------------------------------------------------

# Network parameters
min_jaccard   <- 0.14   # Minimum Jaccard similarity for edge
min_node_n    <- 13     # Minimum document frequency for node
max_node_n    <- 350    # Maximum document frequency for node
label_min_n   <- 13     # Minimum df for label display
degree_filter <- 1      # Minimum degree to keep node

pain_terms <- unigram_lexicon_lemmatized$word

# Filter edges: must involve pain term OR disease term
pain_edges <- pairs_jaccard |>
  filter(
    item1 %in% pain_terms | item2 %in% pain_terms |
    str_detect(item1, disease_pattern) | str_detect(item2, disease_pattern)
  ) |>
  filter(jaccard >= min_jaccard) |>
  transmute(from = item1, to = item2, weight = jaccard)

# Build node list
pain_nodes <- tibble(name = unique(c(pain_edges$from, pain_edges$to))) |>
  left_join(term_doc_freq, by = c("name" = "word")) |>
  mutate(
    df = replace_na(df, 0L),
    is_pain_term = name %in% pain_terms
  ) |>
  filter(df >= min_node_n, df <= max_node_n)

# Filter edges to only include nodes in node list
pain_edges <- pain_edges |>
  filter(from %in% pain_nodes$name & to %in% pain_nodes$name)

# Build graph
g <- tbl_graph(nodes = pain_nodes, edges = pain_edges, directed = FALSE) |>
  activate(nodes) |>
  mutate(deg = centrality_degree()) |>
  filter(deg >= degree_filter)
```

```{r}
#| label: plot-network
#| message: false
#| fig-width: 12
#| fig-height: 8

# -----------------------------------------------------------------------------
# SEMANTIC NETWORK VISUALIZATION
# -----------------------------------------------------------------------------

set.seed(42)

ggraph(g, layout = "stress") +
  geom_edge_link(aes(width = weight), alpha = 0.3) +
  geom_node_point(aes(color = df, shape = is_pain_term), size = 5) +
  geom_node_text(
    aes(label = name, filter = df >= label_min_n),
    repel = TRUE, size = 3
  ) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_color_viridis_c(option = "C", direction = -1) +
  scale_shape_manual(
    name = "Term type",
    values = c(`TRUE` = 17, `FALSE` = 16),
    labels = c(`TRUE` = "Pain term", `FALSE` = "Non-pain term")
  ) +
  guides(
    color = guide_colorbar(order = 1, barheight = unit(35, "mm")),
    shape = guide_legend(order = 2, override.aes = list(color = "#E69F00")),
    edge_width = guide_legend(order = 3)
  ) +
  labs(
    title = "Pain- and Disease-Anchored Semantic Network",
    subtitle = "Top 500 myositis posts by pain centrality",
    caption = "Edges represent normalized co-occurrence (Jaccard similarity).",
    edge_width = "Co-occurrence\nstrength",
    color = "Posts with term"
  ) +
  theme_minimal(base_size = 12) +
  coord_cartesian(clip = "off") +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.position = "right",
    legend.box = "vertical",
    plot.margin = margin(10, 10, 10, 10),
    plot.title.position = "plot",
    plot.caption = element_text(size = 9, color = "grey40", hjust = 0)
  )

cat("Network statistics:\n")
cat("  Nodes:", gorder(g), "\n")
cat("  Edges:", gsize(g), "\n")
```

## 9. Network Centrality Analysis

```{r}
#| label: centrality-measures
#| message: false

# -----------------------------------------------------------------------------
# CENTRALITY MEASURES
# -----------------------------------------------------------------------------

centrality_stats <- g |>
  activate(nodes) |>
  mutate(
    betweenness = centrality_betweenness(),
    degree = centrality_degree()
  ) |>
  as_tibble() |>
  arrange(desc(betweenness)) |>
  slice_head(n = 10)

centrality_table <- centrality_stats |>
  mutate(
    Type = if_else(is_pain_term, "Pain Term", "General/Clinical Term"),
    betweenness = round(betweenness, 1)
  ) |>
  select(
    Term = name,
    Type,
    `Betweenness Centrality` = betweenness,
    `Degree (Connections)` = degree
  )

kable(centrality_table,
      caption = "Top 10 Most Central Terms in the Myositis Pain Network")
```

## 10. Community Detection

```{r}
#| label: community-detection
#| message: false

# -----------------------------------------------------------------------------
# COMMUNITY DETECTION (LOUVAIN ALGORITHM)
# -----------------------------------------------------------------------------

set.seed(123)
gi <- as.igraph(g)
comm <- cluster_louvain(g, weights = E(g)$weight)
V(g)$community <- membership(comm)

# Key descriptives
n_communities <- length(unique(V(g)$community))
Q_modularity <- modularity(comm)

cat("Communities detected:", n_communities, "\n")
cat("Modularity (Q):", round(Q_modularity, 3), "\n")
```

```{r}
#| label: community-summary
#| message: false

# -----------------------------------------------------------------------------
# COMMUNITY COMPOSITION SUMMARY
# -----------------------------------------------------------------------------

community_summary <- tibble(
  term = V(g)$name,
  community = V(g)$community,
  is_pain = V(g)$is_pain_term,
  df = V(g)$df
) |>
  group_by(community) |>
  summarise(
    n_terms = n(),
    pain_terms = sum(is_pain, na.rm = TRUE),
    non_pain_terms = sum(!is_pain, na.rm = TRUE),
    pain_share = round(mean(is_pain, na.rm = TRUE), 2),
    mean_df = round(mean(df, na.rm = TRUE), 1),
    top_terms = paste(term[order(df, decreasing = TRUE)][1:min(25, n())],
                      collapse = ", "),
    .groups = "drop"
  ) |>
  arrange(desc(n_terms))

print(community_summary)

writexl::write_xlsx(community_summary, file.path(DATA_PATH, "data/community_summary_original.xlsx"))
```

```{r}
#| label: community-test
#| message: false

# -----------------------------------------------------------------------------
# STATISTICAL TEST: COMMUNITY COMPOSITION
# -----------------------------------------------------------------------------
# Test whether communities differ in pain vs non-pain term composition

contingency <- table(
  Community = V(g)$community,
  PainTerm = V(g)$is_pain_term
)

# Use Fisher's exact test if expected counts are small
chi <- suppressWarnings(chisq.test(contingency))
if (any(chi$expected < 5)) {
  test <- fisher.test(contingency)
  test_name <- "Fisher's exact test"
} else {
  test <- chi
  test_name <- "Chi-square test"
}

cat("\nCommunity composition test:", test_name, "\n")
print(test)
```

```{r}
#| label: community-plot
#| message: false
#| fig-width: 12
#| fig-height: 8

# -----------------------------------------------------------------------------
# NETWORK COLORED BY COMMUNITY
# -----------------------------------------------------------------------------

ggraph(g, layout = "stress") +
  geom_edge_link(aes(width = weight), alpha = 0.2) +
  geom_node_point(aes(color = factor(community), shape = is_pain_term), size = 4) +
  scale_edge_width(range = c(0.2, 2.5)) +
  scale_shape_manual(
    name = "Term type",
    values = c(`TRUE` = 17, `FALSE` = 16),
    labels = c(`TRUE` = "Pain term", `FALSE` = "Non-pain term")
  ) +
  labs(
    title = "Semantic Network Communities",
    subtitle = paste0("Communities: ", n_communities, " | Modularity: ", round(Q_modularity, 3)),
    edge_width = "Co-occurrence strength",
    color = "Community",
    shape = "Term type"
  ) +
  theme_void()
```

```{r}
#| label: permutation-test
#| message: false

# -----------------------------------------------------------------------------
# PERMUTATION TEST: PAIN TERM DISTRIBUTION
# -----------------------------------------------------------------------------
# Test whether pain terms are clustered within communities more than expected

set.seed(123)
n_perm <- 1000
pain_share_null <- numeric(n_perm)

for (i in seq_len(n_perm)) {
  shuffled <- sample(V(g)$is_pain_term)
  pain_share_null[i] <- tapply(shuffled, V(g)$community, mean) |>
    var(na.rm = TRUE)
}

observed_var <- tapply(V(g)$is_pain_term, V(g)$community, mean) |>
  var(na.rm = TRUE)

p_value_perm <- mean(pain_share_null >= observed_var)

cat("Permutation test (n =", n_perm, "):\n")
cat("  Observed variance in pain share:", round(observed_var, 4), "\n")
cat("  P-value:", round(p_value_perm, 4), "\n")

# Cross-tabulation
cat("\nPain terms by community:\n")
print(table(V(g)$community, V(g)$is_pain_term))
```
