---
title: "Myositis and Pain: Data Import Pipeline"
subtitle: "Importing and merging Reddit data from Pushshift archives and subreddit dumps"
author: "Peter Kerkhof"
date: last-modified
format:
  html: default
  pdf: default
  docx: default
editor: visual
execute:
  echo: false
  warning: false
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---

## Overview

This script imports Reddit submission data from two sources:

1.  **Pushshift Parquet Archives** - Queried from Academic Torrents Reddit dumps
2.  **Subreddit JSONL Archives** - Downloaded from Arctic Shift for specific myositis subreddits

The data is merged, deduplicated, cleaned, and saved as a single feather file for downstream analysis.

------------------------------------------------------------------------

## 1. Setup and Configuration

```{r}
#| message: false
#| label: setup-environment

# -----------------------------------------------------------------------------
# ENVIRONMENT SETUP
# -----------------------------------------------------------------------------
# Clear environment (preserve functions) and reset console
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")

# Disable scientific notation for cleaner number display
options(scipen = 999)
```

```{r}
#| message: false
#| label: load-libraries

# -----------------------------------------------------------------------------
# LOAD REQUIRED LIBRARIES
# -----------------------------------------------------------------------------
library(tidyverse)
library(lubridate)
library(stringr)
library(furrr)
library(jsonlite)
library(here)
library(feather)
library(arrow)

# Load project configuration (DATA_PATH, REDDIT_DATA_PATH)
source(here("config.R"))
```

------------------------------------------------------------------------

## 2. Import Pushshift Query Results

The Pushshift archive contains Reddit data distributed via Academic Torrents. We query this archive for posts containing myositis-related keywords.

```{r}
#| message: false
#| label: import-parquet-query

# -----------------------------------------------------------------------------
# IMPORT PUSHSHIFT PARQUET QUERY RESULTS
# -----------------------------------------------------------------------------

# Define cutoff date: August 1, 2025
# Posts after this date are excluded to ensure a consistent dataset
cut_off_date <- as.numeric(as.POSIXct("2025-08-01", tz = "UTC"))
cut_off_date <- 1754006400 - 7200  # Adjusted for timezone

# To re-run the parquet query (slow), uncomment:
# source(file.path(REDDIT_DATA_PATH, "queries/query_parquet_myositis.R"))

# Load pre-queried results from feather file
# Query searched for: myopath*, myositis, synthetase syndrome, anti-jo-1
# Excluded: cardiomyopathy
df_query <- read_feather(
  file.path(REDDIT_DATA_PATH, "queries/query-output/myositis_submissions20250910.feather")
) |>
  mutate(across(everything(), as.character)) |>
  mutate(source = "parquet") |>
  filter(as.numeric(created_utc) < cut_off_date)
```

------------------------------------------------------------------------

## 3. Import Subreddit Archives

In addition to the keyword query, we import all posts from three myositis-specific subreddits (r/dermatomyositis, r/polymyositis, r/myositis) downloaded from Arctic Shift.

```{r}
#| message: false
#| label: import-subreddit-archives

# -----------------------------------------------------------------------------
# IMPORT SUBREDDIT JSONL ARCHIVES
# -----------------------------------------------------------------------------

# Path to subreddit JSONL files (from Arctic Shift)
folder_path <- file.path(REDDIT_DATA_PATH, "subreddits/data/myositis")

# Find all post JSONL files recursively
files <- list.files(
  path = folder_path,
  pattern = "post.*\\.jsonl$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

# Helper function: Read JSONL file with error handling
read_jsonl <- function(file_path) {
  con <- file(file_path, open = "r")
  tryCatch({
    data <- stream_in(con, verbose = FALSE)
    return(data)
  }, error = function(e) {
    message(sprintf("Error reading %s: %s", file_path, e$message))
    return(NULL)
  }, finally = {
    if (isOpen(con)) close(con)
  })
}

# Read all JSONL files and combine into single dataframe
# Select only columns that exist in the parquet query results
df_subreddits <- map_dfr(files, read_jsonl) |>
  select(any_of(colnames(df_query))) |>
  mutate(across(everything(), as.character)) |>
  mutate(source = "subreddits")

# Store count for reporting
n_posts_subreddits <- nrow(df_subreddits)
```

------------------------------------------------------------------------

## 4. Merge and Deduplicate

```{r}
#| message: false
#| label: merge-and-deduplicate

# -----------------------------------------------------------------------------
# MERGE DATA SOURCES AND REMOVE DUPLICATES
# -----------------------------------------------------------------------------

df <- bind_rows(df_query, df_subreddits) |>
  # Ensure created_utc is numeric for filtering
  mutate(created_utc = as.numeric(created_utc)) |>
  # Apply cutoff date filter
  filter(created_utc < cut_off_date) |>
  # Sort chronologically
  arrange(created_utc) |>
  # Remove duplicates: keep first occurrence (from parquet if both sources have it)
  # Ignore 'source' column when checking for duplicates
  distinct(across(-source), .keep_all = TRUE) |>
  # Convert Unix timestamps to datetime objects
  mutate(
    created_utc = as.POSIXct(as.numeric(created_utc), origin = '1970-01-01'),
    retrieved_on = as.POSIXct(as.numeric(retrieved_on), origin = '1970-01-01')
  ) |>
  # Extract date components for temporal analysis

  mutate(
    date = date(created_utc),
    year = year(created_utc),
    month = month(created_utc),
    month_yr = format_ISO8601(created_utc, precision = "ym"),
    week_number = isoweek(date)
  ) |>
  # Remove comment-specific columns (not relevant for submissions)
  select(-c(body, link_id, parent_id))
```

------------------------------------------------------------------------

## Data Source Summary

Data were retrieved from the Reddit Pushshift archive, distributed via Academic Torrents. Pushshift has collected nearly all publicly available Reddit submissions and comments since 2005 and continues to release monthly archival dumps (Academic Torrents, 2025). Although Reddit revoked Pushshift's real-time API access in 2023, the archival dumps remain continuously updated. We conducted a query for posts containing the words myopath, myositis, synthetase syndrome, and anti-Jo-1 in either the title or the self-text field (the fields that contain the title and the main body of a submission). The search allowed for different variations of these terms (e.g., myopathies, myopathy, dermatomyositis, antijo1, etc.) up to August 1, 2025, while excluding posts containing the word cardiomyopathy. Comments (both to submissions and to other comments) were excluded, as they are less likely to contain patient experiences or to be written with the intent to share a story, seek advice, or narrate symptoms. Submissions are also typically more self-contained: their content can be interpreted without additional conversational context.

The query was deliberately broad, with the resulting data subsequently refined through several cleaning steps. The query was case-insensitive and returned `r scales::comma(nrow(df_query))` posts, including those mentioning myositis and myopath either as standalone words or as parts of longer words (e.g., dermatomyositis, myopathy, myopathies). In addition, we included all submissions (`r n_posts_subreddits`) from three myositis-related subreddits: r/dermatomyositis, r/polymyositis, and r/myositis. In total, the dataset comprised `r scales::comma(nrow(df))` posts published between July 2008 and July 31, 2025.

------------------------------------------------------------------------

## 5. Text Preprocessing

Merge title and selftext into a single text field, with proper punctuation handling.

```{r}
#| message: false
#| label: preprocess-text

# -----------------------------------------------------------------------------
# TEXT PREPROCESSING: Merge title + selftext
# -----------------------------------------------------------------------------

df <- df |>
  mutate(
    # Trim whitespace from text fields
    selftext = str_trim(selftext),
    title = str_trim(title),

    # Check if title already ends with terminal punctuation
    title_punctuated = str_detect(title, "[\\.\\?\\!\\:]$"),

    # Prepare title for merging:
    # - Add ". " if title doesn't end with punctuation
    # - Add " " only if it already has punctuation
    # - Handle NA/empty titles
    title_clean = case_when(
      is.na(title) | title == "" ~ "",
      title_punctuated ~ paste0(title, " "),
      TRUE ~ paste0(title, ". ")
    ),

    # Merge title and selftext into single text field
    text = paste0(title_clean, selftext),

    # Clean up merged text
    text = gsub("^NA\\.\\s*", "", text),  # Remove leading "NA. "
    text = gsub("NA", "", text),           # Remove stray "NA"
    text = tolower(text),                  # Lowercase for consistency
    text = str_replace_all(text, "[-_]", " ")  # Normalize hyphens/underscores
  ) |>
  # Remove temporary columns
  select(-title_punctuated, -title_clean) |>
  # Reorder columns: put key fields first
  select(author, created_utc, text, everything())
```

------------------------------------------------------------------------

## 6. Extract Query Keywords

Identify which search keywords matched in each post for transparency and analysis.

```{r}
#| message: false
#| label: extract-keywords

# -----------------------------------------------------------------------------
# KEYWORD EXTRACTION
# -----------------------------------------------------------------------------
# Track which query keywords matched in each post

# Define search keywords (same as used in parquet query)
keywords <- c(
  "myopath", "myositis", "synthetase syndrome", "synthetasesyndrome",
  "anti-jo-1", "antijo1", "anti jo 1"
)

# Create regex pattern for matching
pattern <- str_c(keywords, collapse = "|")

# Extract all matching keywords from each post
df <- df |>
  mutate(
    keywords_found_list = str_extract_all(text, regex(pattern, ignore_case = TRUE)),
    keywords_found = sapply(keywords_found_list, function(x) {
      if (length(x) == 0 || all(is.na(x))) {
        NA_character_
      } else {
        str_c(unique(x), collapse = ", ")
      }
    })
  ) |>
  select(-keywords_found_list)

# Normalize keyword variations in text
replacements <- c(
  "anti jo 1" = "anti-jo-1",
  "antijo1" = "anti-jo-1"
)
df <- df |>
  mutate(text = str_replace_all(text, replacements))

# Re-extract keywords after normalization (sorted for consistency)
df <- df |>
  mutate(
    keywords_found_list = str_extract_all(text, regex(pattern, ignore_case = TRUE)),
    keywords_found = sapply(keywords_found_list, function(x) {
      if (length(x) == 0 || all(is.na(x))) {
        NA_character_
      } else {
        sorted_unique <- sort(unique(x))
        str_c(sorted_unique, collapse = ", ")
      }
    })
  ) |>
  select(-keywords_found_list)

# Summary table of keyword combinations (for reference)
keyw <- as.data.frame(table(df$keywords_found)) |>
  arrange(desc(Freq))
```

------------------------------------------------------------------------

## 7. Generate Post URLs

Create direct links to each Reddit post for easy reference and verification.

```{r}
#| message: false
#| label: create-post-urls

# -----------------------------------------------------------------------------
# GENERATE POST URLS
# -----------------------------------------------------------------------------

df <- df |>
  mutate(
    post_url = paste0("https://www.reddit.com/comments/", id)
  )
```

------------------------------------------------------------------------

## 8. Author Anonymization

Anonymize author usernames using CRC32 hashing to protect privacy while maintaining the ability to track unique authors.

```{r}
#| message: false
#| label: anonymize-authors

# -----------------------------------------------------------------------------
# ANONYMIZE AUTHOR USERNAMES
# -----------------------------------------------------------------------------

# Calculate word and token counts for each post
df <- df |>
  mutate(
    n_words = str_count(text, "\\w+"),
    n_tokens = quanteda::ntoken(quanteda::tokens(text))  )

# Anonymize author names using CRC32 hash
# Same author always gets same anonymous ID (reproducible with seed)
df$author_ano <- anonymizer::anonymize(
  df$author,
  .algo = "crc32",
  .seed = 0,
  .chars = letters,
  .n_chars = 5L
)

# Replace the hash for "[deleted]" authors with readable label
df <- df |>
  mutate(author_ano = str_replace(author_ano, "3c0a4086", "deleted"))

# Reorder columns
df <- df |>
  select(source, subreddit, created_utc, text, everything())
```

------------------------------------------------------------------------

## 9. Save Output

Export the merged and cleaned dataset to feather format for efficient loading in subsequent analysis scripts.

```{r}
#| message: false
#| label: save-output

# -----------------------------------------------------------------------------
# SAVE PROCESSED DATA
# -----------------------------------------------------------------------------

output_path <- file.path(DATA_PATH, "data", "myositis_submissions_incl_subreddits_2005_july2025.feather")
#write_feather(df, output_path)

message(sprintf("Saved %s rows to: %s", scales::comma(nrow(df)), output_path))
```

------------------------------------------------------------------------

## Output Summary

-   **Total posts**: `r scales::comma(nrow(df))`
-   **From Pushshift query**: `r scales::comma(nrow(df_query))`
-   **From subreddit archives**: `r scales::comma(n_posts_subreddits)`
-   **Date range**: `r min(df$date)` to `r max(df$date)`
-   **Output file**: `DATA_PATH/data/myositis_submissions_incl_subreddits_2005_july2025.feather`
