---
title: "Myositis and Pain: Data Cleaning Pipeline"
subtitle: "Cleaning and preprocessing Reddit data for myositis analysis"
author: "Peter Kerkhof"
date: last-modified
format:
  html: default
  pdf: default
  docx: default
editor: visual
execute:
  echo: false
  warning: false
editor_options:
  chunk_output_type: inline
always_allow_html: yes
---

## Overview

This script cleans Reddit submission data for myositis-related analysis. The cleaning pipeline removes irrelevant, low-quality, and non-English posts through a series of filtering steps, each documented with case counts.

## 1. Setup and Data Loading

```{r}
#| message: false
#| label: setup-environment

# -----------------------------------------------------------------------------
# ENVIRONMENT SETUP
# -----------------------------------------------------------------------------
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")

options(scipen = 999)
```

```{r}
#| message: false
#| label: load-libraries

# -----------------------------------------------------------------------------
# LOAD REQUIRED LIBRARIES
# -----------------------------------------------------------------------------
library(tidyverse)
library(lubridate)
library(here)
library(furrr)
library(jsonlite)
library(arrow)
library(feather)
library(udpipe)
library(tidytext)
library(stringi)
library(cld2)
library(purrr)
library(knitr)
library(kableExtra)

# Load project configuration (DATA_PATH)
source(here("config.R"))
```

```{r}
#| message: false
#| label: load-data

# -----------------------------------------------------------------------------
# LOAD RAW DATA
# -----------------------------------------------------------------------------
df_start <- arrow::read_feather(
  file.path(DATA_PATH, "data", "myositis_submissions_incl_subreddits_2005_july2025.feather")
)

df <- df_start

# -----------------------------------------------------------------------------
# INITIALIZE CASE TRACKING
# -----------------------------------------------------------------------------
# This tibble will track how many cases are removed at each cleaning step
case_flow <- tibble(
  step = character(),
  description = character(),
  n_before = integer(),
  n_removed = integer(),
  n_after = integer()
)

# Helper function to log cleaning steps
log_step <- function(step_name, description, n_before, n_after) {
  case_flow <<- case_flow |>
    add_row(
      step = step_name,
      description = description,
      n_before = n_before,
      n_removed = n_before - n_after,
      n_after = n_after
    )
}

# Log initial state
n_initial <- nrow(df_start)
```

**Initial Dataset:** Before cleaning, the data contained `r scales::comma(nrow(df_start))` Reddit posts, posted by `r scales::comma(length(unique(df_start$author)))` unique authors across `r scales::comma(length(unique(df_start$subreddit)))` subreddits.

**Cleaning Steps Summary:**

The following posts were removed during cleaning:

-   Posts containing only generic "myopathy/ies" keywords (not inflammatory/necrotizing)
-   Posts about non-inflammatory myositis types (tension, orbital, ossificans, pyomyositis)
-   Duplicate posts (identical text within 14 days by same author)
-   Posts with deleted/removed content
-   Posts from bot accounts
-   Non-English posts
-   Posts with excessive URLs (≥5) or asterisks (≥11)
-   Posts with fewer than 10 words
-   Posts containing promotional content ("product name")
-   Posts from animal-related or otherwise irrelevant subreddits
-   Posts from personal subreddits (u/)

## 2. Keyword Filtering: Myositis-Related Terms

This step ensures posts specifically reference inflammatory or necrotizing myopathies, not generic muscle conditions.

```{r}
#| message: false
#| label: filter-myositis-keywords

# -----------------------------------------------------------------------------
# STEP 2A: FILTER FOR SPECIFIC MYOSITIS-RELATED KEYWORDS
# -----------------------------------------------------------------------------
# Posts containing only generic "myopathy/ies" are excluded unless they
# also mention inflammatory/necrotizing variants or other myositis terms.

n_before <- nrow(df)

# Define acceptable myositis-related keywords
# These terms specifically indicate inflammatory/autoimmune muscle disease
myo_keywords <- c(

  "necrotizing myopathy", "necrotizing myopathies",
  "necrotising myopathy", "necrotising myopathies",
  "inflammatory myopathy", "inflammatory myopathies",
  "myositis", "polymyositis", "dermatomyositis",
  "synthetase syndrome", "synthetasesyndrome",
  "anti-jo-1", "antijo1", "anti jo 1"
)

# Escape special characters and allow suffix variations (e.g., plural forms)
myo_keywords <- str_replace_all(myo_keywords, "[-()]", "\\\\$0")
myo_keywords <- paste0(myo_keywords, "\\w*")
keyword_pattern <- paste(myo_keywords, collapse = "|")

# Identify posts that would be excluded (for reporting)
df_excluded_myopath <- df |>
  filter(
    !(
      str_detect(text, regex(keyword_pattern, ignore_case = TRUE)) |
      str_detect(subreddit, regex(keyword_pattern, ignore_case = TRUE))
    )
  )
n_excluded_myopath <- nrow(df_excluded_myopath)

# Keep only posts with myositis-related keywords in text OR subreddit name
df <- df |>
  filter(
    str_detect(text, regex(keyword_pattern, ignore_case = TRUE)) |
    str_detect(subreddit, regex(keyword_pattern, ignore_case = TRUE))
  )

n_after_keywords <- nrow(df)
log_step("2a", "Generic myopathy keywords (not inflammatory/necrotizing)", n_before, n_after_keywords)
```

```{r}
#| message: false
#| label: filter-non-inflammatory-myositis

# -----------------------------------------------------------------------------
# STEP 2B: EXCLUDE NON-INFLAMMATORY MYOSITIS TYPES
# -----------------------------------------------------------------------------
# Remove posts about myositis conditions outside the study scope:
# - Tension myositis syndrome (psychosomatic condition)
# - Pyomyositis (bacterial infection)
# - Myositis ossificans (bone formation in muscle)
# - Orbital myositis (eye muscle inflammation)

n_before <- nrow(df)

df <- df |>
  filter(
    # Exclude tension myositis (syndrome)
    !str_detect(
      paste(subreddit, text, sep = " "),
      regex("tension\\s*myositis(?:syndrome)?", ignore_case = TRUE)
    ) &
    # Exclude pyomyositis (bacterial infection of muscle)
    !str_detect(text, regex("pyomyositis", ignore_case = TRUE)) &
    # Exclude myositis ossificans (heterotopic bone formation)
    !str_detect(text, regex("myositis ossificans", ignore_case = TRUE)) &
    # Exclude orbital myositis (eye muscle inflammation)
    !str_detect(text, regex("orbital myositis", ignore_case = TRUE))
  )

n_after <- nrow(df)
log_step("2b", "Non-inflammatory myositis types (tension, orbital, ossificans, pyomyositis)", n_before, n_after)

cat("Step 2 complete - Keyword filtering\n")
cat("  Rows after step 2:", scales::comma(nrow(df)), "\n")
```

**Step 2 Results:** Since "myopathy" is a broad term that also refers to non-inflammatory muscle conditions, we excluded posts that did not explicitly reference necrotizing or inflammatory myopathy/ies (n = `r scales::comma(n_excluded_myopath)`). Posts about tension myositis, orbital myositis, myositis ossificans, and pyomyositis—all outside the scope of this study—were also excluded (n = `r scales::comma(n_before - n_after)`), leaving `r scales::comma(nrow(df))` posts.

## 3. Subreddit and Author Filtering

Remove posts from irrelevant subreddits (e.g., animal health, unrelated topics) and known bot accounts.

```{r}
#| label: filter-subreddits-authors
#| message: false

# -----------------------------------------------------------------------------
# STEP 3: FILTER SUBREDDITS AND AUTHORS
# -----------------------------------------------------------------------------

n_before <- nrow(df)

# Define excluded subreddits: general irrelevant topics
exclude_values_general <- c(
  # Academic/study-related
  "Step1", "Step2", "scholar", "paidstudies", "PaidStudyHub", "paidstudy",
  "medicalschool", "medicalschoolanki", "anki", "NBME",
  # News/entertainment aggregators
  "autotldr", "noveltranslations", "TheDahmerCase", "Dahmer",
  "skulduggerypleasant", "thesopranos", "funny", "OnePiece", "lakers",
  # Conspiracy/misinformation

  "conspiracy", "conspiracyNOPOL", "conspiracyundone", "DebateVaccine", "VaxTalk",
  # Finance/investment
  "CABAinvest", "wallstreetbets", "weedstocks",
  # Technology
  "buildapc", "Watches", "remotedaily", "ChinaInnovation",
  # Health conditions unrelated to myositis
  "Cardiomyopathy", "Vitiligo", "PSSD", "COVIDZero",
  # Miscellaneous irrelevant
  "sarasota", "googletranslate", "takecareofmayaFree", "motivation",
  "zmarter", "ketoscience", "FuckTravisScott", "Autoimmunefaq",
  "SupplementsGuide", "tabled", "MongoloidCJ", "PackItUpinfo",
  "dirtypenpals", "manywords"
)

# Define excluded subreddits: animal-related
# Animals can develop myositis, but these posts are not relevant to human patient experiences
exclude_values_animals <- c(
  # Veterinary/general pet
  "AskVet", "vet", "veterinary", "Veterinary", "VetTech", "veterinarian",
  "Pets", "Petloss", "PetAdvice", "petyoloblog", "National_Pet_Adoption",
  # Dogs (general and breeds)
  "dogs", "puppy101", "Greyhounds", "CaneCorso", "GreatDane", "goldenretrievers",
  "roughcollies", "BelgianMalinois", "germanshepherds", "pitbulls", "Boxer",
  "DobermanPinscher", "greatdanes", "Catahoula", "BorderCollie", "BorderTerrier",
  "BostonTerrier", "BoxerDogs", "Boxers", "Chihuahua", "akita", "corgi",
  "Pitbull", "greatpyrenees", "ridgebacks", "bernesemountaindog",
  "bernesemountaindogs", "u_bobbiesocksthepug",
  # Cats
  "cats", "CatAdvice", "CATHELP", "blackcats", "bengalcats", "catfood",
  "SiberianCats", "standardissuecat", "catcare", "RenalCats", "Siamesecats",
  "TuxedoCats", "IllegallySmolCats", "Feral_Cats", "catpics", "SupermodelCats",
  "britishshorthair", "BritishShorthaircats", "bengalcatfacts", "bengals",
  # Other animals
  "Rabbits", "Horses", "Equestrian", "AdviceAnimals", "aww",
  "BeardedDragons", "BeardedDragonHelp", "BeardedDragon",
  "BirdHealth", "AfricanGrey"
)

# Define excluded authors: known bots and irrelevant accounts
unwanted_authors <- c(
  "myatomsareyouratoms", "Asher_potter", "D_Upadhaya", "Mentioned_Videos",
  "hibernating_brain", "Mk_Makanaki", "subredditreports", "-NervousPudding-",
  "0100001001010011", "Firm-Twist-4497", "MuscularDystrophy-ModTeam",
  "Rabbits-ModTeam", "SlightlyMadAngus", "Lupinepublishers-ACR",
  "mushroomlover345", "Gir0", "FrozenPixel", "namenerd400"
)

# Combine and lowercase excluded subreddits
excluded_subreddits <- tolower(c(exclude_values_general, exclude_values_animals))

# Apply filters
df <- df |>
  mutate(
    subreddit_lc = tolower(subreddit),
    title0 = coalesce(title, ""),
    selftext0 = coalesce(selftext, "")
  ) |>
  filter(
    !(
      # Exclude listed subreddits
      subreddit_lc %in% excluded_subreddits |
      # Exclude penny stock subreddits
      str_detect(subreddit, regex("penny", ignore_case = TRUE)) |
      # Exclude dog-related subreddits not in list
      str_detect(subreddit, regex("dog", ignore_case = TRUE)) |
      # Exclude personal subreddits (u_username format)
      str_detect(subreddit, regex("^u_", ignore_case = TRUE)) |
      # Exclude unwanted authors
      author %in% unwanted_authors |
      # Exclude posts about masticatory myositis (dog condition)
      str_detect(title0, regex("masticatory", ignore_case = TRUE)) |
      str_detect(selftext0, regex("masticatory", ignore_case = TRUE))
    )
  ) |>
  select(-subreddit_lc, -title0, -selftext0)

# Remove bot accounts based on username patterns
df <- df |>
  filter(!str_detect(author, regex("bot$", ignore_case = TRUE))) |>
  filter(!str_detect(author, regex("^auto", ignore_case = TRUE))) |>
  filter(!str_detect(author, regex("admin$", ignore_case = TRUE)))

# Remove posts mentioning Jeffrey Dahmer (serial killer who had polymyositis)
df <- df |>
  filter(!str_detect(text, regex("dahmer", ignore_case = TRUE)))

n_after <- nrow(df)
log_step("3", "Irrelevant subreddits, authors, bots, and off-topic content", n_before, n_after)

cat("Step 3 complete - Subreddit/author filtering\n")
cat("  Rows after step 3:", scales::comma(nrow(df)), "\n")
```

**Step 3 Results:** Posts from animal-related subreddits (animals can develop myositis), clearly irrelevant subreddits (e.g., about serial killer Jeffrey Dahmer, a polymyositis patient), personal subreddits (u/), and bot accounts were removed. This excluded `r scales::comma(n_before - n_after)` posts, leaving `r scales::comma(nrow(df))` posts.

## 4. Remove Deleted Content and Duplicates

Remove posts with deleted/removed content and near-duplicate posts by the same author.

```{r}
#| message: false
#| label: filter-deleted-duplicates

# -----------------------------------------------------------------------------
# STEP 4: REMOVE DELETED CONTENT AND DUPLICATES
# -----------------------------------------------------------------------------

n_before <- nrow(df)

# Remove posts where title or selftext is empty, [removed], or [deleted]
df <- df |>
  mutate(
    across(c(title, selftext), ~ str_trim(.x)),
    across(c(title, selftext), ~ na_if(.x, ""))
  ) |>
  filter(
    !if_any(
      c(title, selftext),
      ~ is.na(.x) |
        str_detect(.x, regex("^\\s*\\[(removed|deleted)\\]\\s*$", ignore_case = TRUE))
    )
  )

n_after_deleted <- nrow(df)
n_deleted <- n_before - n_after_deleted

# Flag and remove duplicates within 14 days by same author
# Duplicates are identified by identical title OR selftext content
df <- df |>
  mutate(
    created_utc = as.POSIXct(created_utc, origin = "1970-01-01"),
    rowid = row_number()
  )

# Helper function to flag duplicates for a given column
dup_flag <- function(dat, col) {
  col_sym <- sym(col)

  dat |>
    filter(!is.na(!!col_sym)) |>
    group_by(author, !!col_sym) |>
    arrange(created_utc, .by_group = TRUE) |>
    mutate(
      prev = lag(created_utc),
      .dup = !is.na(prev) & as.numeric(difftime(created_utc, prev, units = "days")) <= 14
    ) |>
    ungroup() |>
    transmute(rowid, !!paste0(col, "_dup") := .dup)
}

# Apply duplicate flagging to title and selftext columns
cols <- intersect(c("title", "selftext"), names(df))
flags <- purrr::map(cols, ~ dup_flag(df, .x)) |>
  reduce(full_join, by = "rowid")

# Remove flagged duplicates (keep earliest post in each duplicate set)
df <- df |>
  left_join(flags, by = "rowid") |>
  mutate(
    across(ends_with("_dup"), ~ coalesce(.x, FALSE)),
    to_remove = if_any(ends_with("_dup"), identity)
  ) |>
  filter(!to_remove) |>
  select(-rowid, -ends_with("_dup"), -to_remove)

n_after <- nrow(df)
n_duplicates <- n_after_deleted - n_after
log_step("4", "Deleted/removed content and duplicate posts", n_before, n_after)

cat("Step 4 complete - Deleted content and duplicate removal\n")
cat("  Deleted/removed content:", scales::comma(n_deleted), "\n")
cat("  Duplicate posts:", scales::comma(n_duplicates), "\n")
cat("  Rows after step 4:", scales::comma(nrow(df)), "\n")
```

**Step 4 Results:** Posts with empty, \[removed\], or \[deleted\] title/selftext were removed (n = `r scales::comma(n_deleted)`). Potential duplicates—posts with identical title or selftext by the same author within 14 days—were also removed, keeping only the earliest occurrence (n = `r scales::comma(n_duplicates)`). This left `r scales::comma(nrow(df))` posts.

## 5. Remove Spam and Low-Quality Content

Filter out posts with excessive formatting markers, URLs, or list-like structure that lack coherent narrative.

```{r}
#| message: false
#| label: filter-spam-formatting

# -----------------------------------------------------------------------------
# STEP 5: REMOVE SPAM AND LOW-QUALITY CONTENT
# -----------------------------------------------------------------------------

n_before <- nrow(df)

# Count URLs and asterisks in title and selftext
df <- df |>
  mutate(
    # Count URLs (http://, https://, www.)
    n_urls_title = str_count(title, pattern = "https?://\\S+|www\\.\\S+"),
    n_urls_selftext = str_count(selftext, pattern = "https?://\\S+|www\\.\\S+"),
    # Count asterisks (often used in spam/formatting)
    n_asterisks_title = str_count(title, fixed("*")),
    n_asterisks_selftext = str_count(selftext, fixed("*"))
  ) |>
  replace_na(list(
    n_urls_title = 0, n_urls_selftext = 0,
    n_asterisks_title = 0, n_asterisks_selftext = 0
  )) |>
  mutate(
    n_urls_total = n_urls_title + n_urls_selftext,
    n_asterisks_total = n_asterisks_title + n_asterisks_selftext
  )

# Remove posts with excessive asterisks (≥11), URLs (≥5), or "product name"
df <- df |>
  filter(
    n_asterisks_total < 11,
    n_urls_total < 5,
    is.na(selftext) | !str_detect(selftext, regex("product name", ignore_case = TRUE))
  )

# Identify and remove list-like posts
# These are posts dominated by short lines (e.g., symptom lists, supplement lists)
# that don't provide interpretable patient narratives
df <- df |>
  mutate(
    # Count line breaks
    n_lines = str_count(selftext, "\n"),
    # Split into lines
    lines = str_split(selftext, "\n"),
    # Count words per line
    line_lengths = map(lines, ~ str_count(.x, "\\S+")),
    # Calculate proportion of lines with < 5 words
    prop_short = map_dbl(line_lengths, ~ mean(.x < 5, na.rm = TRUE)),
    # Flag: list-like if ≥20 lines AND ≥70% are short
    is_list_like = n_lines >= 20 & prop_short >= 0.7
  )

df <- df |> filter(!is_list_like)

# Also filter posts that are 100% short sentences unless they have ≥50 lines
df <- df |>
  filter(prop_short != 1 | n_lines >= 50)

n_after <- nrow(df)
log_step("5", "Spam, excessive formatting, and list-like content", n_before, n_after)

cat("Step 5 complete - Spam/formatting removal\n")
cat("  Rows after step 5:", scales::comma(nrow(df)), "\n")
```

**Step 5 Results:** Posts containing excessive URLs (≥5), formatting markers (≥11 asterisks), or explicit promotional cues ("product name") were removed. Posts dominated by long sequences of short, list-like lines (e.g., symptom or supplement lists) that lack coherent narrative structure were also excluded. This step removed `r scales::comma(n_before - n_after)` posts, leaving `r scales::comma(nrow(df))` posts.

## 6. Language and Length Filtering

Remove non-English posts and posts with insufficient content.

```{r}
#| message: false
#| label: filter-language-length

# -----------------------------------------------------------------------------
# STEP 6: FILTER BY LANGUAGE AND MINIMUM LENGTH
# -----------------------------------------------------------------------------

n_before <- nrow(df)

df <- df |>
  filter(
    # Require minimum 10 words for meaningful content
    n_words > 9,
    # Keep only English-language posts (using cld2 language detection)
    cld2::detect_language(text) == "en",
    # Remove known non-English subreddit
    !is.na(subreddit),
    subreddit != "omoshiroi_eigo_bunpou"
  )

n_after <- nrow(df)
log_step("6", "Non-English posts and posts with < 10 words", n_before, n_after)

cat("Step 6 complete - Language/length filtering\n")
cat("  Rows after step 6:", scales::comma(nrow(df)), "\n")
```

**Step 6 Results:** Non-English posts and posts with fewer than 10 words were excluded, leaving `r scales::comma(nrow(df))` posts from `r scales::comma(length(unique(df$author)))` unique authors.

## 7. Case Flow Summary

```{r}
#| message: false
#| label: case-flow-summary
#| echo: false

# Add final row to case flow
case_flow <- case_flow |>
  add_row(
    step = "Final",
    description = "Clean dataset",
    n_before = nrow(df),
    n_removed = 0,
    n_after = nrow(df)
  )

# Calculate percentages
case_flow <- case_flow |>
  mutate(
    pct_removed = round(n_removed / n_initial * 100, 1),
    pct_remaining = round(n_after / n_initial * 100, 1)
  )

# Display case flow table
case_flow |>
  select(Step = step, Description = description,
         `N Before` = n_before, `N Removed` = n_removed, `N After` = n_after,
         `% Removed` = pct_removed, `% Remaining` = pct_remaining) |>
  kable(format.args = list(big.mark = ",")) |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

**Summary:** The cleaning pipeline reduced the dataset from `r scales::comma(n_initial)` to `r scales::comma(nrow(df))` posts (`r round(nrow(df)/n_initial*100, 1)`% retained).

## 8. Text Preprocessing and Lemmatization

Create a cleaned and lemmatized version of the text for downstream analysis.

```{r}
#| message: false
#| label: lemmatization

# -----------------------------------------------------------------------------
# TEXT PREPROCESSING AND LEMMATIZATION
# -----------------------------------------------------------------------------

# Download and load English udpipe model
download_info <- udpipe_download_model(language = "english-ewt")
model_path_downloaded <- download_info$file_model

if (!file.exists(model_path_downloaded)) {
  stop("Failed to download the English udpipe model.")
}

ud_model <- udpipe_load_model(file = model_path_downloaded)

# Light cleaning + Unicode normalization (keep punctuation for UDPipe)
df$text_clean <- df$text |>
  str_replace_all("https?://\\S+|www\\.\\S+", " ") |>
  stri_trans_general("NFKC") |>
  str_replace_all("\\p{Cf}", " ") |>
  str_squish()

# Ensure doc_id column exists
if (!"doc_id" %in% names(df)) {
  df$doc_id <- as.character(seq_len(nrow(df)))
}

# Annotate text with udpipe
x <- udpipe_annotate(ud_model, x = df$text_clean, doc_id = df$doc_id)
x_df <- as.data.frame(x) |>
  mutate(lemma = ifelse(tolower(token) == "feedback", "feedback", lemma))

# Post-process: keep content words; fix common lemmatization errors
content_upos <- c("NOUN", "PROPN", "ADJ", "VERB")

# Domain-specific term mappings to protect key medical terms
domain_map <- c(
  "joint" = "joint", "joints" = "joint",
  "flare" = "flare", "flares" = "flare",
  "ra" = "ra", "ccp" = "ccp", "rf" = "rf",
  "myositis" = "myositis", "neuropathy" = "neuropathy"
)

lemmas_filtered <- x_df |>
  transmute(
    doc_id, token = tolower(token), lemma = tolower(lemma), upos
  ) |>
  # Fix bad POS tags for long alphabetic tokens
  mutate(
    bad_tag = (!upos %in% content_upos) & grepl("^[a-z][a-z'-]{3,}$", token),
    lemma = if_else(bad_tag, token, lemma)
  ) |>
  # Fall back to token if lemma has weird characters
  mutate(lemma = if_else(!grepl("^[a-z][a-z'-]*$", lemma), token, lemma)) |>
  # Apply domain dictionary overrides
  mutate(lemma = if_else(token %in% names(domain_map), domain_map[token], lemma)) |>
  # Keep only content words (nouns, proper nouns, adjectives, verbs)
  filter(upos %in% content_upos) |>
  # Remove stopwords (but keep "back" when used as noun)
  filter((lemma == "back" & upos == "NOUN") | !(lemma %in% stop_words$word)) |>
  # Remove common contractions that slip through
  filter(!(lemma %in% c("ive", "im", "dont", "cant", "wont"))) |>
  # Fix udpipe error: "joint" → "jous"
  mutate(lemma = if_else(lemma == "jous", "joint", lemma))

# Rebuild lemmatized text
lemmatized_text_df <- lemmas_filtered |>
  group_by(doc_id) |>
  summarise(lemmatized_text = paste(lemma, collapse = " "), .groups = "drop")

df <- df |> left_join(lemmatized_text_df, by = "doc_id")
```

## 9. Save Cleaned Data

```{r}
#| message: false
#| label: save-output

# -----------------------------------------------------------------------------
# SAVE CLEANED DATA
# -----------------------------------------------------------------------------

# Uncomment to save:
# write_feather(df, file.path(DATA_PATH, "data", "myositis_submissions_2005_july2025_clean.feather"))
# write_feather(x_df, file.path(DATA_PATH, "data", "myositis_lemmatized_udpipe.feather"))

cat("Cleaning complete.\n")
cat("Final dataset:", scales::comma(nrow(df)), "posts from",
    scales::comma(length(unique(df$author))), "unique authors\n")
```

## 10. Descriptive Statistics

### Posts Over Time

```{r}
#| message: false
#| label: plot-posts-over-time

df_summary <- df |>
  mutate(year_month = as.Date(floor_date(created_utc, "month"))) |>
  group_by(year_month) |>
  summarize(count = n())

ggplot(df_summary, aes(x = year_month, y = count)) +
  geom_point() +
  geom_line(alpha = 0.3) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(
    title = "Myositis-Related Reddit Posts per Month",
    x = "Year",
    y = "Number of Posts"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The number of myositis-related posts has grown substantially, reaching nearly 100 per month in 2024, with steep growth after 2020 (mirroring Reddit's overall growth).

### Posts by Year

```{r}
#| message: false
#| label: table-posts-by-year

df |>
  count(year) |>
  rename(freq = n) |>
  mutate(percentage = round((freq / sum(freq)) * 100, 1)) |>
  arrange(desc(year)) |>
  mutate(cum_percentage = round(cumsum(percentage), 1)) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Over half of submissions were posted in 2024 or 2025.

### Top Subreddits

```{r}
#| message: false
#| label: table-top-subreddits

df |>
  count(subreddit) |>
  rename(freq = n) |>
  mutate(percentage = round((freq / sum(freq)) * 100, 1)) |>
  arrange(desc(freq)) |>
  mutate(cum_percentage = round(cumsum(percentage), 1)) |>
  slice_head(n = 10) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Myositis mentions span many subreddits, including dedicated communities (r/dermatomyositis) and general health forums (r/AskDocs, r/Autoimmune).

### Myositis Keyword Frequencies

```{r}
#| message: false
#| label: table-keyword-frequencies

df$text <- paste(df$title, df$selftext, sep = " ")
text_col <- tolower(df$text)
text_col <- gsub("_", " ", text_col)

# Match words containing "myositis"
pattern <- "\\b\\w*myositis\\w*\\b"
matches <- str_extract_all(text_col, pattern)
matched_words <- unlist(matches)
word_counts <- table(matched_words)
word_counts_df <- as.data.frame(word_counts)
colnames(word_counts_df) <- c("Word", "Count")

word_counts_df |>
  mutate(percentage = round((Count / sum(Count)) * 100, 1)) |>
  arrange(desc(Count)) |>
  mutate(cum_percentage = round(cumsum(percentage), 1)) |>
  slice_head(n = 20) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

### Posts with Personal Pronouns

```{r}
#| message: false
#| label: pronoun-analysis

pronoun_pattern <- "\\b(i|me|my|she|he|her|his)\\b"

df_pronouns <- df |>
  filter(str_detect(text, regex(pronoun_pattern, ignore_case = TRUE)))

df_no_pronouns <- df |>
  filter(!str_detect(text, regex(pronoun_pattern, ignore_case = TRUE)))

cat("Posts with personal pronouns:", scales::comma(nrow(df_pronouns)),
    "(", round(nrow(df_pronouns)/nrow(df)*100, 1), "%)\n")
cat("Posts without personal pronouns:", scales::comma(nrow(df_no_pronouns)),
    "(", round(nrow(df_no_pronouns)/nrow(df)*100, 1), "%)\n")
```
