---
title: "Myositis"
format:
  html: default
  pdf: default
  docx: default
editor: visual
execute:
  echo: false
  warning: false
editor_options: 
  chunk_output_type: inline
always_allow_html: yes
---

## Importing the data

```{r}
#| message: false
#| label: loading-libraries-and-data

rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
cat("\014")
options(scipen = 999) 


# Loading required libraries
library(tidyverse); library(lubridate)
library(stringr); library(furrr); library(jsonlite); library(here);library(feather)
library(arrow)

source(here("config.R"))

# Uncomment to rerun parquet data extraction
#source(file.path(REDDIT_DATA_PATH, "queries/query_parquet_myositis.R"))

cut_off_date <- as.numeric(as.POSIXct("2025-08-01", tz = "UTC"))
cut_off_date <- 1754006400 - 7200

df_query <- read_feather(file.path(REDDIT_DATA_PATH, "queries/query-output/myositis_submissions20250910.feather")
) |>
  mutate(across(everything(), as.character)) |>
  mutate(source = "parquet") |>
  filter(as.numeric(created_utc) < cut_off_date) 

folder_path <- file.path(REDDIT_DATA_PATH,"subreddits/data/myositis")

# List all JSONL files in the /data directory
files <- list.files(
  path = folder_path,
  pattern = "post.*\\.jsonl$",
  full.names = TRUE,
  recursive = TRUE,
  ignore.case = TRUE
)

read_jsonl <- function(file_path) {
  con <- file(file_path, open = "r") # Open the connection
  tryCatch({
    data <- stream_in(con, verbose = FALSE)
    return(data)
  }, error = function(e) {
    message(sprintf("Error reading %s: %s", file_path, e$message))
    return(NULL)
  }, finally = {
    if (isOpen(con)) {
      close(con)
    }
  })
}

df_subreddits <- map_dfr(files, read_jsonl) |>
  select(any_of(colnames(df_query))) |>
  mutate(across(everything(), as.character)) |>
  mutate(source = "subreddits") 

df <- bind_rows(df_query, df_subreddits) |>
  mutate(created_utc = as.numeric(created_utc)) |>  # fix hier
  filter(created_utc < cut_off_date)  |>
  arrange(created_utc) |>
  distinct(across(-source), .keep_all = TRUE) |>
  mutate(
    created_utc = as.POSIXct(as.numeric(created_utc), origin = '1970-01-01'),
    retrieved_on = as.POSIXct(as.numeric(retrieved_on), origin = '1970-01-01')
  ) |>
  mutate(
    date = date(created_utc),
    year = year(created_utc),
    month = month(created_utc),
    month_yr = format_ISO8601(created_utc, precision = "ym"),
    week_number = isoweek(date)
  )|>
  select(c(-body,-link_id, -parent_id))

n_posts_subreddits <- nrow(df_subreddits) 

```

Data were retrieved from the reddit Pushshift archive, distributed via Academic Torrents. Pushshift has collected nearly all publicly available Reddit submissions and comments since 2005 and continues to release monthly archival dumps (Academic Torrents, 2025). Although Reddit revoked Pushshift’s real-time API access in 2023, the archival dumps remain continuously updated. We conducted a query for posts containing the words myopath, myositis, synthetase syndrome, and anti-Jo-1 in either the title or the self-text field (the fields that contain the title and the main body of a submission). The search allowed for different variations of these terms (e.g., myopathies, myopathy, dermatomyositis, antijo1, etc.) up to August 1, 2025, while excluding posts containing the word cardiomyopathy. Comments (both to submissions and to other comments) were excluded, as they are less likely to contain patient experiences or to be written with the intent to share a story, seek advice, or narrate symptoms. Submissions are also typically more self-contained: their content can be interpreted without additional conversational context.

The query was deliberately broad, with the resulting data subsequently refined through several cleaning steps. The query was case-insensitive and returned `r scales::comma(nrow(df_query))` posts, including those mentioning myositis and myopath either as standalone words or as parts of longer words (e.g., dermatomyositis, myopathy, myopathies). In addition, we included all submissions (`r n_posts_subreddits`) from three myositis-related subreddits: r/dermatomyositis, r/polymyositis, and r/myositis. In total, the dataset comprised `r scales::comma(nrow(df))` posts published between July 2008 and July 31, 2025.

```{r}
#| message: false
#| label: trim and merge all text columns; adding a dot after the title

#rm(list = setdiff(ls(), "df"))

df <- df |>
  mutate(
    selftext = str_trim(selftext),
    title    = str_trim(title),

    # Determine whether title ends with punctuation: . ? ! :
    title_punctuated = str_detect(title, "[\\.\\?\\!\\:]$"),

    # Add ". " only if:
    #  - title is not NA/empty
    #  - AND it does not already end with punctuation
    title_clean = case_when(
      is.na(title) | title == "" ~ "",
      title_punctuated ~ paste0(title, " "),        # ends with punctuation → add space only
      TRUE ~ paste0(title, ". ")                    # no punctuation → add ". "
    ),

    # Merge title + selftext
    text = paste0(title_clean, selftext),

    # Cleanup
    text = gsub("^NA\\.\\s*", "", text),
    text = gsub("NA", "", text),
    text = tolower(text),
    text = str_replace_all(text, "[-_]", " ")
  ) |>
  select(-title_punctuated, -title_clean) |>
  select(author, created_utc, text, everything())

```

```{r}
#| message: false
#| label: establishing which query keyword

# Define your keywords
keywords <- c(
  "myopath", "myositis", "synthetase syndrome", "synthetasesyndrome", "anti-jo-1", "antijo1", "anti jo 1"
)

# Create a single regex pattern for case-insensitive matching
pattern <- str_c(keywords, collapse = "|")

# Extract matched keywords from the combined text column
df <- df |>
  mutate(
    keywords_found_list = str_extract_all(text, regex(pattern, ignore_case = TRUE)),
    keywords_found = sapply(keywords_found_list, function(x) {
      if (length(x) == 0 || all(is.na(x))) {
        NA_character_
      } else {
        str_c(unique(x), collapse = ", ")
      }
    })
  ) |>
  select(-keywords_found_list)

# Define replacements
replacements <- c(
  "anti jo 1" = "anti-jo-1",
  "antijo1"   = "anti-jo-1")

# Apply replacements
df <- df |> 
  mutate(text = str_replace_all(text, replacements))

#remove different outcomes based on the order of keywords
df <- df |>
  mutate(
    keywords_found_list = str_extract_all(text, regex(pattern, ignore_case = TRUE)),
    keywords_found = sapply(keywords_found_list, function(x) {
      if (length(x) == 0 || all(is.na(x))) {
        NA_character_
      } else {
        sorted_unique <- sort(unique(x))
        str_c(sorted_unique, collapse = ", ")
      }
    })
  ) |>
  select(-keywords_found_list)

keyw <- as.data.frame(table(df$keywords_found)) |>
  arrange(desc(Freq))
```

```{r}
#| message: false
#| label: create post url

df <- df |>
  mutate(
    post_url = paste0(
      "https://www.reddit.com/comments/",
      id
    )
  )

```

```{r}
#| message: false
#| label: create anonymous user accounts

df <- df |>
  mutate(
    n_words = str_count(text, "\\w+"),
    n_tokens = ntoken(tokens(text)),
  )

#anonymize authors
df$author_ano <- anonymizer::anonymize(df$author, 
                                         .algo = "crc32", .seed = 0, 
                                         .chars = letters,
                                         .n_chars = 5L)
df <-df |> #rename deleted author names
  mutate(author_ano = str_replace(author_ano, "3c0a4086", "deleted")) 

df <- df |>
  select(source, subreddit, created_utc, text, everything()) 

```

```{r}
#| message: false
#| label: determine post type



```{r}
#| message: false
#| label: create post type column

# Create post_type column to distinguish submissions, comments, and comments to comments
df <- df |>
  mutate(
    post_type = case_when(
      is.na(parent_id) | parent_id == "" ~ "submission",
      str_starts(parent_id, "t3_") ~ "comment",
      str_starts(parent_id, "t1_") ~ "comment_to_comment",
      TRUE ~ "unknown"
    )
  )

# Show distribution of post types
post_type_counts <- df |>
  count(post_type, sort = TRUE) |>
  mutate(
    percentage = scales::percent(n / sum(n), accuracy = 0.1)
  )

print(post_type_counts)

write_feather(df, file.path(DATA_PATH, "data", "myositis_submissions_incl_subreddits_2005_july2025.feather"))

```




